% Attempt for Erdos Problem #9
% Following PROMPT_STRATEGY.MD
% Tools/Constraints:
% - Web browsing available? NO (restricted)
% - Computation available? YES (local Python), not used
% Codex 5.2 extra high

OUTPUT
1) FORMAL RESTATEMENT
Define
\[A:=\{n\in 2\mathbb Z+1: \text{$n$ is not of the form } p+2^k+2^l\text{ with }k,l\ge 0,\ p\text{ prime}\}.\]
Question: Is the upper density
\[\overline d(A):=\limsup_{N\to\infty} \frac{|A\cap\{1,\dots,N\}|}{N}
\]
positive?

2) QUICK LITERATURE/CONTEXT CHECK
Browsing is not available. I only record what the problem statement itself claims.
I have not verified those results here.

3) ATTACK PLAN
Proof track:
1. Use sieve methods to show representable numbers have density < 1 among odd integers.
2. Control the union over all pairs $(k,l)$ of translated primes.

Disproof track:
1. Try to show representable numbers are dense by covering odd integers with $p+2^k+2^l$.

Chosen path: establish elementary counting bounds and a baseline construction of representable odds.

4) WORK
Lemma 1 (Infinitely many representable odd integers).
There are infinitely many odd integers of the form $p+2^k+2^l$.

Proof.
Fix any $k,l\ge 0$. As $p$ ranges over odd primes, $p+2^k+2^l$ ranges over infinitely many odd integers. \qed

Lemma 2 (Crude upper bound on the count of representable integers up to $N$).
Let $R(N)$ be the number of integers $\le N$ that are representable as $p+2^k+2^l$. Then
\[R(N)\le \pi(N)\cdot (1+\lfloor \log_2 N\rfloor)^2,\]
where $\pi(N)$ is the prime-counting function.

Proof.
For each pair $(k,l)$ with $2^k+2^l\le N$, the count of primes $p$ with $p+2^k+2^l\le N$ is at most $\pi(N)$.
There are at most $(1+\lfloor \log_2 N\rfloor)^2$ such pairs. Summing over all pairs gives the bound. \qed

Lemma 3 (A trivial inclusion for $p=2$).
Every integer of the form $2+2^k+2^l$ is representable.

Proof.
Take $p=2$ in the defining form. \qed

5) VERIFICATION
- Lemma 2 is a union bound; it does not attempt to estimate overlaps.
- Lemma 3 is purely definitional.

6) FINAL
**UNRESOLVED**

(i) Strongest fully proved partial result:
A crude counting upper bound: $R(N)\le \pi(N)(1+\lfloor\log_2 N\rfloor)^2$.

(ii) First gap:
No nontrivial lower bound on $|A\cap[1,N]|$ is proved; in particular, no positive-density bound.

(iii) Top 3 next moves:
1. Use sieve methods to estimate the density of $p+2^k+2^l$ uniformly in $(k,l)$.
2. Study distribution of $2^k+2^l$ modulo small primes to create local obstructions.
3. Attempt a lower bound on $|A\cap[1,N]|$ via a large-sieve inequality.

(iv) Minimal counterexample structure (if the conjecture were false):
A scenario where the union of shifted primes $\bigcup_{k,l}(\mathcal P+2^k+2^l)$ has density 1 within the odd integers,
requiring highly uniform prime distribution across the shifts by all sums of two powers of 2.
