% Erdos Problem #1041
% URL: https://www.erdosproblems.com/1041

Let $f(z)=\prod_{i=1}^n(z-z_i)\in \mathbb{C}[z]$ with $\lvert z_i\rvert < 1$ for all $i$. Must there always exist a path of length less than $2$ in\[\{z: \lvert f(z)\rvert < 1\}\]which connects two of the roots of $f$? A problem of Erd\H{o}s, Herzog, and Piranian \cite{EHP58}, who proved that this set always has a connected component containing at least two of the roots. References [EHP58] Erd\H{o}s, P. and Herzog, F. and Piranian, G., Metric properties of polynomials . J. Analyse Math. (1958), 125-148.


% --- SOLUTION TO ERDOS PROBLEM #1041 ---

FORMAL RESTATEMENT
Let f(z)=\prod_{i=1}^n (z-z_i) with all roots satisfying |z_i|<1, and let
  L(f) := { z in C : |f(z)| < 1 }.
Question: must there exist indices i!=j and a rectifiable path gamma in L(f) connecting z_i to z_j such that the Euclidean length of gamma is strictly less than 2?
(For n=1 the question is vacuous; the interesting case is n>=2.)

QUICK LITERATURE/CONTEXT CHECK
Only facts explicitly present in the problem statement are treated as established here. In particular, it is stated that L(f) always has a connected component containing at least two roots. The additional length <2 requirement is the new constraint.

ATTACK PLAN
(1) Prove the statement for n=2 exactly (this is the smallest nontrivial case).
(2) Find a clean sufficient condition guaranteeing that the straight segment between two roots stays inside L(f).
(3) Attempt to argue that such a sufficient condition must occur for some pair, or else try to build counterexamples by spreading roots.

WORK
FAST REALITY CHECK.
* n=2 is fully solved below: there is always a path of length |z_1-z_2|<2 (Lemma 1041.1).
* For n=3,4,5,6 a random experiment (roots uniform in the unit disc) shows that the sufficient condition in Lemma 1041.2 holds for at least one pair with frequencies:
  n=3: about 0.641
  n=4: about 0.619
  n=5: about 0.540
  n=6: about 0.448
This does not prove the conjecture but is a sanity check that the condition is not extremely rare for small n.

Lemma 1041.1 (the conjecture holds for n=2).
Let f(z)=(z-a)(z-b) with |a|<1 and |b|<1 and a!=b. Then there is a path in L(f) from a to b of length |a-b|<2.

Proof.
Let d:=|a-b|. Since |a|,|b|<1, the triangle inequality gives d <= |a|+|b| < 2, hence d<2.
Let m=(a+b)/2 be the midpoint.
Consider the straight segment from a to m, parameterized by
  z(t) = a + t(m-a) = a + t(b-a)/2,  0<=t<=1.
Then |z(t)-a| = t d/2.
Also,
  z(t)-b = a-b + t(b-a)/2 = -(1-t/2)(b-a),
so |z(t)-b| = (1-t/2) d.
Therefore, along this segment we have
  |f(z(t))| = |z(t)-a||z(t)-b| = (t d/2) * ((1-t/2)d) = (t(1-t/2)) * d^2/2.
For t in [0,1], the function t(1-t/2) is increasing (its derivative is 1-t>=0), so its maximum on [0,1] is attained at t=1 and equals 1/2.
Hence for all t in [0,1],
  |f(z(t))| <= (1/2) * d^2/2 = d^2/4 < 1.
Thus the entire segment from a to m lies in L(f).
By symmetry, the same argument shows the segment from m to b also lies in L(f).
Concatenating these two segments gives a path from a to b inside L(f) of total length
  |a-m| + |m-b| = d/2 + d/2 = d < 2.

Lemma 1041.2 (a sufficient condition: close pair implies a short connecting segment).
Let n>=2 and f(z)=\prod_{k=1}^n (z-z_k) with all |z_k|<1. Suppose there exist i!=j with d:=|z_i-z_j| such that
  d^2 (2+d)^{n-2} < 1.
Then the straight line segment [z_i,z_j] is contained in L(f). In particular, z_i and z_j can be connected inside L(f) by a path of length d (<2).

Proof.
Let z be any point on the straight segment from z_i to z_j. Then by convexity,
  |z-z_i| <= d  and  |z-z_j| <= d.
Now fix any other root z_k with k not in {i,j}. Using the triangle inequality,
  |z-z_k| <= |z-z_i| + |z_i-z_k|.
Since |z_i|<1 and |z_k|<1, we have |z_i-z_k| <= |z_i|+|z_k| < 2. Thus |z-z_k| < d+2.
Combining these bounds over all factors,
  |f(z)| = \prod_{k=1}^n |z-z_k| < d * d * (2+d)^{n-2} = d^2(2+d)^{n-2} < 1.
Therefore every z on [z_i,z_j] lies in L(f), and the segment length is d<2.

VERIFICATION
* Lemma 1041.1 is a direct computation along a segment and is fully rigorous.
* Lemma 1041.2 is a direct product bound using only triangle inequality and the root constraints.
* The random experiment numbers were computed by a minimal script; they are not used as proof.

FINAL
UNRESOLVED
(i) Strongest proved partial result here: the statement is true for n=2 (Lemma 1041.1). More generally, if two roots are sufficiently close in the sense d^2(2+d)^{n-2}<1, then the straight segment between them lies in {|f|<1} and yields a path of length <2 (Lemma 1041.2).
(ii) First gap: show that for every n>=2 and every choice of roots with |z_i|<1, there exists some pair i!=j satisfying a condition strong enough to force a path of length <2 inside L(f); or else construct an explicit counterexample polynomial.
(iii) Top 3 next moves:
  1. Try to strengthen Lemma 1041.2 by replacing the crude bound |z-z_k|<2+d with something depending on where the segment lies (e.g. if the segment stays well inside the unit disc, improve the 2 constant).
  2. Combine a packing argument (some pair of roots must be close) with a lower bound on how small the other factors can be along a connecting corridor; attempt to force |f|<1 along a short path.
  3. Computational: for small degrees (say n<=6) attempt an explicit search over root configurations to maximize the minimal in-domain path length between roots (e.g. discretize the plane and do shortest-path computations in the sublevel set) to look for counterexamples or patterns.
(iv) Minimal counterexample structure (if false): roots arranged near the unit circle and spread so that (a) the component containing multiple roots exists but (b) any connecting path must detour around regions where |f|>=1, forcing length >=2.

