% Erdos Problem #1112

1) FORMAL RESTATEMENT

Fix integers \(1\le d_1<d_2\) and \(k\ge 3\).  Let \(r\ge 2\) be an integer.

A (strictly) increasing sequence of positive integers \(B=\{b_1<b_2<\cdots\}\) is called \(r\)-lacunary if
\[
 b_{i+1}\ge r\,b_i\quad\text{for all }i\ge 1.
\]

A (strictly) increasing sequence of positive integers \(A=\{a_1<a_2<\cdots\}\) has \((d_1,d_2)\)-bounded gaps if
\[
 d_1\le a_{i+1}-a_i\le d_2\quad\text{for all }i\ge 1.
\]

Define the \(k\)-fold sumset (with repetition allowed)
\[
 kA := \underbrace{A+\cdots +A}_{k\text{ times}}=\{a_{i_1}+\cdots+a_{i_k}: i_1,\dots,i_k\ge 1\}.
\]

Question (as interpreted in the problem statement): does there exist an integer \(r=r_k(d_1,d_2)\) such that for every \(r\)-lacunary \(B\) there exists an infinite \(A\) with \((d_1,d_2)\)-bounded gaps and
\[
 (kA)\cap B = \varnothing\ ?
\]
(Edge case note: if one required the sumset to use \emph{distinct} elements of \(A\), the problem changes; the text ``\(k\)-fold sumset'' is standardly interpreted as allowing repetition, and the computations/lemmas below use that.)

2) QUICK LITERATURE/CONTEXT CHECK

I did not use web browsing here; I only record what is explicitly stated in the problem text:

- For \(d_1=2,d_2=3\), Erd\H{o}s--Graham noted the existence of \(A\) with gaps in \([2,3]\) such that \((A+A)\cap B=\emptyset\) for any \(B\) with \(b_{1}\ge 5\) and \(b_{i+1}\ge 2b_i\).
- Bollob\'as--Hegyv\'ari--Jin proved sharp results for \(k=2\) in that regime and also that \(r_3(2,3)\) does \emph{not} exist (in the sense stated in the problem text).
- Chen proved \(r_2(a,b)\le 2\) for \(b\ne 2a\) and lower bounds for \(b=2a\).
- For \(k\ge 3\) beyond the nonexistence examples, the general existence question remains open, with further technical nonexistence results by Tang--Yang.

3) ATTACK PLAN

Proof strategies (if aiming for existence of \(r_k(d_1,d_2)\)):

- Try to build \(A\) as a ``near-periodic'' bounded-gap sequence and modify it locally near each forbidden value \(b_i\), using the lacunarity to keep modifications far apart.
- Try modular obstructions: force \(kA\) to land in a controlled set of residues mod \(q\), then choose \(q\) so that \(B\) avoids those residues.

Disproof/construction strategies (to show \(r_k(d_1,d_2)\) does not exist):

- Construct a universal \(r\)-lacunary set \(B\) that intersects \(kA\) for \emph{every} \((d_1,d_2)\)-bounded-gap \(A\). (This is what the cited nonexistence results accomplish in special cases.)

Below I only obtain basic structural lemmas and a few explicit examples; they do not settle the general existence/nonexistence problem.

4) WORK

\textbf{Lemma 1112.1 (Counting bounds for bounded-gap \(A\)).}
Let \(A=\{a_1<a_2<\cdots\}\) have \((d_1,d_2)\)-bounded gaps.  Then for every integer \(N\ge a_1\), writing \(A(N):=|A\cap [a_1,N]|\), we have
\[
 1+\frac{N-a_1}{d_2} \le A(N) \le 1+\frac{N-a_1}{d_1}.
\]
In particular the lower asymptotic density of \(A\) is at least \(1/d_2\).

\emph{Proof.}
For the upper bound: if \(A(N)=t\), then \(a_t\le N\). Summing the lower gap bound,
\(a_t-a_1=\sum_{i=1}^{t-1}(a_{i+1}-a_i)\ge (t-1)d_1\), hence \(N-a_1\ge (t-1)d_1\), i.e.
\(t\le 1+(N-a_1)/d_1\).

For the lower bound: let \(t\) be maximal with \(a_t\le N\) (so \(A(N)=t\)). Then
\(a_{t+1}>N\). Using the upper gap bound,
\(a_{t+1}\le a_1+(t)d_2\) because
\(a_{t+1}-a_1=\sum_{i=1}^{t}(a_{i+1}-a_i)\le t d_2\).
Thus \(N<a_{t+1}\le a_1+t d_2\), so \(t> (N-a_1)/d_2\), i.e.
\(t\ge 1+(N-a_1)/d_2\) since \(t\) is an integer. \qed

\textbf{Lemma 1112.2 (A uniform syndeticity bound for \(kA\) from the upper gap bound).}
Let \(A=\{a_1<a_2<\cdots\}\) be an increasing sequence of integers with \(a_{i+1}-a_i\le d_2\) for all \(i\) (no lower gap assumption needed). Fix \(k\ge 1\) and set
\(M:=k(d_2-1)\).
Then every integer \(n\ge k a_1\) belongs to
\[
 kA + \{0,1,\dots,M\}.
\]
Equivalently, for every integer \(n\ge k a_1\) there exists \(s\in kA\) with \(n-M\le s\le n\). In particular, the set \(kA\) has gaps at most \(M+1\) beyond \(k a_1\).

\emph{Proof.}
First observe that the condition \(a_{i+1}-a_i\le d_2\) implies
\[
\{m\in\mathbb{Z}: m\ge a_1\}\subseteq A+\{0,1,\dots,d_2-1\}.
\]
Indeed, let \(m\ge a_1\), and choose \(i\) maximal with \(a_i\le m\). Then \(m<a_{i+1}\), hence
\(0\le m-a_i\le a_{i+1}-a_i-1\le d_2-1\), so \(m=a_i+t\) with \(t\in\{0,\dots,d_2-1\}\).

Now take \(n\ge k a_1\) and write
\(n = x_1+\cdots+x_k\) where \(x_1=\cdots=x_{k-1}=a_1\) and \(x_k:=n-(k-1)a_1\ge a_1\).
Each \(x_j\) lies in \(A+\{0,\dots,d_2-1\}\) by the previous paragraph. Thus we may write
\(x_j=a_{i_j}+t_j\) with \(a_{i_j}\in A\) and \(t_j\in\{0,\dots,d_2-1\}\). Summing,
\[
 n = (a_{i_1}+\cdots+a_{i_k}) + (t_1+\cdots+t_k),
\]
where the first parenthesis lies in \(kA\) and the second lies in \(\{0,1,\dots,k(d_2-1)\}=\{0,1,\dots,M\}\). This proves \(n\in kA+\{0,\dots,M\}\).

Finally, if a set \(S\subset\mathbb{Z}\) satisfies \([N,\infty)\subseteq S+\{0,\dots,M\}\), then every interval \([t,t+M]\) with \(t\ge N\) must intersect \(S\); otherwise no element of \([t,t+M]\) could be written as \(s+u\) with \(s\in S\), \(0\le u\le M\). This implies the gaps of \(S\) beyond \(N\) are at most \(M+1\). Apply this with \(S=kA\) and \(N=k a_1\). \qed

\textbf{Lemma 1112.3 (A parity-based explicit avoidance example for odd \(k\)).}
Assume \(2\in[d_1,d_2]\) and \(k\) is odd. Let
\(A:=\{1,3,5,\dots\}\). Then \(A\) has \((d_1,d_2)\)-bounded gaps (all gaps equal to \(2\)), and every element of \(kA\) is odd. Consequently, \(kA\cap B=\emptyset\) for every \(B\subseteq 2\mathbb{Z}\) (in particular for \(B=\{2^j:j\ge 1\}\)).

\emph{Proof.}
All gaps in \(A\) equal \(2\), so if \(2\in[d_1,d_2]\) then \(d_1\le a_{i+1}-a_i\le d_2\) holds.
If \(k\) is odd, the sum of \(k\) odd integers is odd, so \(kA\subseteq 2\mathbb{Z}+1\). Hence \(kA\) is disjoint from any subset of the even integers. \qed

\textbf{FAST REALITY CHECK (one concrete instance).}
Take \((d_1,d_2,k)=(2,3,3)\) and the lacunary set \(B=\{2^j:j\ge 0\}\) (ratio \(2\)). Lemma 1112.3 gives \(A=\{1,3,5,\dots\}\), and then \(3A\) consists only of odd integers \(\ge 3\), while \(B\) consists of \(1\) and even integers, so \((3A)\cap B=\emptyset\).

(Computation note: a backtracking script was run for this instance; it immediately found the same solution \(A\) consisting of consecutive odd integers, which is consistent with the parity argument above.)

5) VERIFICATION

- Lemma 1112.1: the inequalities are obtained by summing gap bounds; all steps are explicit.
- Lemma 1112.2 crucially uses the convention that \(kA\) allows repetition (we used \(a_1\) multiple times).
- Lemma 1112.3: check that the endpoint \(1\in A\), and the smallest element of \(kA\) is \(k\), so \(1\notin kA\) when \(k\ge 3\), ensuring disjointness from \(B\) even if \(1\in B\).

6) FINAL

**UNRESOLVED**

(i) Strongest proved partial result: For any increasing \(A\) with gaps \(\le d_2\), the sumset \(kA\) is syndetic with explicit gap bound \(\le k(d_2-1)+1\) beyond \(k a_1\) (Lemma 1112.2). Also, for odd \(k\) and when \(2\in[d_1,d_2]\), there is an explicit \(A\) with \((d_1,d_2)\)-bounded gaps whose \(k\)-fold sumset avoids any even lacunary \(B\) (Lemma 1112.3).

(ii) First gap (crisp): Given arbitrary integers \(1\le d_1<d_2\) and \(k\ge 3\), prove or disprove the existence of a universal integer \(r\) such that \emph{every} \(r\)-lacunary \(B\) can be avoided by some \((d_1,d_2)\)-bounded-gap \(A\) in the sense \((kA)\cap B=\emptyset\).

(iii) Top 3 next moves:
1. Try to build \(A\) by a lacunary-stage construction: keep \(A\) periodic (fixed step pattern) on long blocks, and at stage \(i\) locally modify a bounded number of steps inside a window around \(b_i\) to eliminate all representations of \(b_i\) as a \(k\)-sum, using the largeness of \(b_{i+1}/b_i\) to keep windows disjoint.
2. For fixed \((d_1,d_2,k)\), do systematic computer search for small \(r\) and specific lacunary \(B\) (e.g. \(b_i=r^i\) or random lacunary) to see whether greedy/local-modification heuristics succeed or repeatedly get trapped.
3. Attempt to extract a structural obstruction from Lemma 1112.2: since \(kA\) is syndetic with an explicit gap bound, look for constructions of \(B\) that force intersection with every syndetic set of that gap type \emph{that is representable as a \(k\)-fold sumset of a \((d_1,d_2)\)-bounded-gap set}.

(iv) Minimal counterexample structure (if \(r_k(d_1,d_2)\) fails to exist): one expects a lacunary \(B\) with extremely fast growth but chosen adversarially so that each \(b_i\) is forced to lie in \(kA\) for every \((d_1,d_2)\)-bounded-gap \(A\); such a \(B\) would likely be constructed inductively to hit the (apparently large) unavoidable subset of \(kA\) common to all admissible \(A\).


