%Erdos problem #156
\section*{Erdos problem \#156}

\subsection*{1) FORMAL RESTATEMENT}
A Sidon set $A\subseteq\{1,2,\dots,N\}$ is \emph{maximal} (with respect to inclusion) if there is no $x\in\{1,\dots,N\}\setminus A$ such that $A\cup\{x\}$ is still Sidon.
The question asks whether there always exists a maximal Sidon set of size $O(N^{1/3})$.

\subsection*{2) QUICK LITERATURE/CONTEXT CHECK}
The problem file states that it is easy to show a greedy construction produces a maximal Sidon set of size $\gg N^{1/3}$, and (as stated in the file) Ruzsa proved existence of maximal Sidon sets of size $\ll (N\log N)^{1/3}$. I do not use those results as black boxes below; I re-prove the $\Omega(N^{1/3})$ lower bound for maximality.

\subsection*{3) ATTACK PLAN}
A maximal Sidon set $A$ blocks every $x\notin A$ by forcing a sum-collision if $x$ were added. Those collisions imply that every $x$ lies in a structured set built from $A$ such as $A+A-A$ or $(A+A)/2$. Counting the size of these structured sets gives a lower bound on $|A|$.

\subsection*{4) WORK}
\paragraph{Lemma 156.1 (maximality forces a covering by $A$, $(A+A)/2$, and $A+A-A$).}
Let $A\subseteq\{1,\dots,N\}$ be Sidon and maximal. Then every $x\in\{1,\dots,N\}$ satisfies at least one of:
\[
\text{(i) }x\in A,\qquad
\text{(ii) }2x\in A+A,\qquad
\text{(iii) }x\in A+A-A.
\]
Equivalently,
\[
\{1,\dots,N\}\subseteq A\ \cup\ \tfrac12(A+A)\ \cup\ \(A+A-A\).
\]

\textit{Proof.}
Fix $x\in\{1,\dots,N\}\setminus A$. By maximality, $A\cup\{x\}$ is not Sidon, so among the new unordered sums created by adding $x$ there is a collision with an old sum.
The new sums are of two forms:
\begin{enumerate}
\item $x+x=2x$,
\item $x+a$ for $a\in A$.
\end{enumerate}
A collision means either:
\begin{enumerate}
\item[(a)] $2x = a+b$ for some $a,b\in A$ (possibly $a=b$). Then $2x\in A+A$, i.e. $x\in\tfrac12(A+A)$.
\item[(b)] $x+a = b+c$ for some $a,b,c\in A$. Rearranging gives $x=b+c-a\in A+A-A$.
\end{enumerate}
Thus every $x\notin A$ lies in $\tfrac12(A+A)$ or in $A+A-A$, establishing the containment. \qed

\paragraph{Lemma 156.2 (a crude but explicit lower bound $|A|=\Omega(N^{1/3})$ for maximal Sidon sets).}
Let $A\subseteq\{1,\dots,N\}$ be maximal Sidon and let $m=|A|$. Then
\[
N \le m + m^2 + m^3,
\]
and hence for all $N\ge 1$,
\[
|A|\ge \bigl(N/3\bigr)^{1/3}.
\]

\textit{Proof.}
From Lemma 156.1,
\[
N\le |A| + |\tfrac12(A+A)| + |A+A-A|.
\]
Trivially $|\tfrac12(A+A)|\le |A+A|\le m^2$.
Also $|A+A-A|\le m^3$ because it is the image of $A^3$ under $(b,c,a)\mapsto b+c-a$.
Thus $N\le m + m^2 + m^3 \le 3m^3$ for $m\ge 1$, which implies $m\ge (N/3)^{1/3}$. \qed

\paragraph{FAST REALITY CHECK (computed small cases).}
I brute-forced the minimum size of a maximal Sidon set in $\{1,\dots,N\}$ for $N\le 40$.
Let $m_{\min}(N)$ be this minimum.
For selected values:
\[
\begin{array}{c|c|l}
N & m_{\min}(N) & \text{example maximal Sidon set}\\\hline
10 & 3 & \{1,4,6\}\\
20 & 4 & \{1,4,10,12\}\\
30 & 5 & \{1,4,10,12,30\}\\
40 & 5 & \{1,4,10,12,30\}
\end{array}
\]
The full sequence for $2\le N\le 40$ is:
\[
2,2,3,3,3,3,3,3,3,4,4,4,4,4,4,4,4,4,4,4,4,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5,5.
\]
(Starting at $N=2$.) This is compatible with growth slower than $N^{1/3}$ in this tiny range, but not decisive.

\subsection*{5) VERIFICATION}
\begin{itemize}
\item Lemma 156.1: enumerates exactly the new sums created by adding $x$ and uses maximality to force a collision.
\item Lemma 156.2: uses only cardinality bounds $|A+A|\le m^2$ and $|A+A-A|\le m^3$.
\item Computation: checked by exhaustive backtracking over Sidon subsets and testing maximality by attempted extensions.
\end{itemize}

\subsection*{6) FINAL}
\textbf{UNRESOLVED}

\begin{enumerate}
\item[(i)] \textbf{Strongest fully proved partial result obtained here.}
Any maximal Sidon set $A\subseteq\{1,\dots,N\}$ has size $|A|\ge (N/3)^{1/3}$ (Lemma 156.2).

\item[(ii)] \textbf{Exact first gap.}
Either (a) construct, for every $N$, a maximal Sidon set of size $\le C N^{1/3}$ with an absolute constant $C$, or (b) prove that every maximal Sidon set has size $\omega(N^{1/3})$.

\item[(iii)] \textbf{Top 3 next moves (concrete targets).}
\begin{enumerate}
\item Sharpen Lemma 156.2 by bounding $|A+A-A|$ substantially below $m^3$ for Sidon sets (Sidon structure should reduce collisions).
\item Try explicit constructions built from ``blocks'' that are Sidon internally but whose cross-block sums are arranged to enforce maximality.
\item Extend brute computations to $N\approx 100$ to guess whether $m_{\min}(N)$ behaves like $c N^{1/3}$.
\end{enumerate}

\item[(iv)] \textbf{Minimal counterexample structure (if $O(N^{1/3})$ is false).}
A counterexample would be a sequence $N\to\infty$ such that every maximal Sidon set in $\{1,\dots,N\}$ has size at least $N^{1/3}\,g(N)$ with $g(N)\to\infty$. By Lemma 156.1, this would force the covering $A\cup \tfrac12(A+A)\cup(A+A-A)$ to be ``inefficient'' unless $|A|$ is larger than $N^{1/3}$ by a diverging factor.
\end{enumerate}


