
PROBLEM

A finite set $A\subset \mathbb{R}^n$ is called Ramsey if, for any $k\geq 1$, there exists some $d=d(A,k)$ such that in any $k$-colouring of $\mathbb{R}^d$ there exists a monochromatic copy of $A$. Characterise the Ramsey sets in $\mathbb{R}^n$.

OUTPUT

## 1) FORMAL RESTATEMENT

Fix $n\ge 1$.
A finite set $A\subseteq\mathbb R^n$ is called **Ramsey** if for every integer $k\ge 1$ there exists an integer $d=d(A,k)\ge 1$ such that for every coloring
\[
c:\mathbb R^d\to\{1,2,\dots,k\}
\]
there exists an isometric embedding (distance-preserving map)
\[
\varphi:\mathbb R^n\to\mathbb R^d
\]
and a color $\gamma$ such that $c$ is constant equal to $\gamma$ on $\varphi(A)$.

The problem asks for a characterization of such finite sets $A$.

---

## 2) QUICK LITERATURE/CONTEXT CHECK

I will not use or claim any external results beyond what is explicitly written in the problem text.

The problem text reports:

* Every Ramsey set is spherical (lies on a sphere).
* Two conjectures: spherical $\Rightarrow$ Ramsey (Graham), and Ramsey $\Leftrightarrow$ subtransitive (Leader–Russell–Walters).
* Several families known to be Ramsey (rectangles, simplices, trapezoids, regular polygons/polyhedra).

I do not re-prove these deep results.

---

## 3) ATTACK PLAN

A full characterization is beyond my current reach. I instead prove basic closure properties that any proposed characterization must respect.

---

## 4) WORK

### Lemma 4.1 (congruence invariance)

If $A\subseteq\mathbb R^n$ is Ramsey and $A'$ is congruent to $A$ (i.e. $A'=\psi(A)$ for some isometry $\psi$ of $\mathbb R^n$), then $A'$ is Ramsey.

**Proof.**
Fix $k\ge 1$.
Since $A$ is Ramsey, there exists $d=d(A,k)$ such that every $k$-coloring of $\mathbb R^d$ contains a monochromatic copy of $A$.

Let $c:\mathbb R^d\to\{1,\dots,k\}$ be any $k$-coloring.
By the Ramsey property of $A$, there exists an isometric embedding $\varphi:\mathbb R^n\to\mathbb R^d$ such that $c$ is constant on $\varphi(A)$.

Now consider $A'=\psi(A)$. The composition $\varphi\circ \psi^{-1}$ is also an isometric embedding of $\mathbb R^n$ into $\mathbb R^d$, and
\[
(\varphi\circ\psi^{-1})(A')=(\varphi\circ\psi^{-1})(\psi(A))=\varphi(A).
\]
Therefore $c$ is constant on $(\varphi\circ\psi^{-1})(A')$, giving a monochromatic copy of $A'$.
Since $c$ was arbitrary, the same $d$ works for $A'$.
$\square$

---

### Lemma 4.2 (heredity under taking subsets)

If $A\subseteq\mathbb R^n$ is Ramsey and $B\subseteq A$ is a (finite) subset, then $B$ is Ramsey.

**Proof.**
Fix $k\ge 1$ and let $d=d(A,k)$ witness that $A$ is Ramsey for $k$ colors.
Given any $k$-coloring $c$ of $\mathbb R^d$, there exists an isometric embedding $\varphi$ such that $c$ is constant on $\varphi(A)$.
Then in particular $c$ is constant on $\varphi(B)\subseteq \varphi(A)$, so $\varphi(B)$ is a monochromatic copy of $B$.
Thus the same $d$ works for $B$.
$\square$

---

## 5) VERIFICATION

* Lemma 4.1: the only step is that composition with an isometry preserves distances, so it maps congruent copies to congruent copies.
* Lemma 4.2: immediate from the definition; any monochromatic copy of $A$ contains a monochromatic copy of any subset.

These lemmas are correct but do not approach the deep classification problem.

---

## 6) FINAL

**UNRESOLVED**

(i) **Strongest fully proved partial result obtained here.**

I proved two structural necessities for the class of Ramsey sets:

* invariance under congruence (Lemma 4.1);
* heredity under taking subsets (Lemma 4.2).

(ii) **Exact first gap.**

Give a nontrivial necessary-and-sufficient geometric characterization of Ramsey sets beyond these basic closure properties.

(iii) **Top 3 next moves (concrete targets).**

1. Prove “spherical $\Rightarrow$ Ramsey” or find a spherical counterexample.
2. Understand the “subtransitive” condition and either prove it implies Ramsey or find a Ramsey set that is not subtransitive.
3. Develop obstructions: show that certain non-spherical configurations cannot be Ramsey by constructing explicit colorings avoiding them.

(iv) **What a minimal counterexample would likely look like.**

A minimal counterexample to “spherical $\Rightarrow$ Ramsey” would be a spherical finite set $A$ with a family of colorings in increasing dimension avoiding monochromatic copies. Such a counterexample would likely be highly symmetric (to evade standard Ramsey embeddings) yet fail the subtransitivity property.


