\section*{Problem \#242 (Erd\H{o}s--Straus Conjecture)}

\subsection*{FORMAL RESTATEMENT}
For each integer $n>2$, determine whether there exist \emph{distinct} positive integers $x<y<z$ such that
\[
\frac{4}{n}=\frac{1}{x}+\frac{1}{y}+\frac{1}{z}.
\]

\subsection*{QUICK LITERATURE/CONTEXT CHECK}
The prompt notes this is the classical Erd\H{o}s--Straus conjecture. The greedy algorithm always gives an Egyptian fraction expansion, but not necessarily with exactly three terms.

\subsection*{ATTACK PLAN}
\textbf{Proof track.}
\begin{itemize}[leftmargin=2.2em]
\item Prove explicit parametrized solutions for infinite families of $n$ (e.g. $n$ even, or $3\mid n$).
\item Prove reduction lemmas (e.g. from composites to primes).
\item Use these to shrink the ``hard'' set of $n$.
\end{itemize}
\textbf{Disproof track.}
\begin{itemize}[leftmargin=2.2em]
\item Search for counterexamples computationally in the remaining residue classes (typically $n\equiv 1,5\pmod 6$).
\item Try to show any putative counterexample must satisfy strong congruence restrictions.
\end{itemize}

\subsection*{WORK}

\begin{lemma}[Scaling]
If $\dfrac{4}{n}=\dfrac1x+\dfrac1y+\dfrac1z$ with $x,y,z\in\mathbb{N}$ distinct, then for any $m\in\mathbb{N}$,
\[
\frac{4}{mn}=\frac1{mx}+\frac1{my}+\frac1{mz}
\]
with $mx,my,mz$ distinct.
\end{lemma}

\begin{proof}
Divide the original identity by $m$.
\end{proof}

\begin{corollary}[Reduction to primes]
If the conjecture fails for some $n>2$, then it fails for some prime $p>2$.
\end{corollary}

\begin{proof}
If $n$ is composite, write $n=mp$ with $p$ prime. If the conjecture holds for $p$, then by Lemma~2 it holds for $mp=n$. Hence any counterexample must have all prime factors failing; choosing one prime factor yields a prime counterexample.
\end{proof}


\begin{proposition}[All even $n$]
If $n>2$ is even, then there exist distinct integers $x<y<z$ with $\frac{4}{n}=\frac1x+\frac1y+\frac1z$.
\end{proposition}

\begin{proof}
Write $n=2m$ with $m\ge 2$. Set
\[
 x:=m,\qquad y:=m+1,\qquad z:=m(m+1).
\]
Then $x<y<z$ for $m\ge 2$, and
\[
\frac1y+\frac1z=\frac{1}{m+1}+\frac{1}{m(m+1)}=\frac{m+1}{m(m+1)}=\frac1m.
\]
Hence $\frac1x+\frac1y+\frac1z=\frac1m+\frac1m=\frac{2}{m}=\frac{4}{n}$.
\end{proof}


\begin{proposition}[All $n$ divisible by $3$]
If $3\mid n$ and $n>2$, then there exist distinct integers $x<y<z$ with $\frac{4}{n}=\frac1x+\frac1y+\frac1z$.
\end{proposition}

\begin{proof}
Write $n=3m$ with $m\ge 1$. Set
\[
 x:=m,\qquad y:=3m+1,\qquad z:=3m(3m+1).
\]
Then $x<y<z$ for $m\ge 1$. Moreover,
\[
\frac1y+\frac1z=\frac{1}{3m+1}+\frac{1}{3m(3m+1)}=\frac{3m+1}{3m(3m+1)}=\frac1{3m}.
\]
Thus $\frac1x+\frac1y+\frac1z=\frac1m+\frac1{3m}=\frac{4}{3m}=\frac{4}{n}$.
\end{proof}

\begin{corollary}[A large solved set of $n$]
The conjectured representation holds for every $n>2$ with $\gcd(n,6)>1$ (i.e. $2\mid n$ or $3\mid n$).
\end{corollary}

\begin{proof}
If $2\mid n$ use Proposition~4; if $3\mid n$ use Proposition~5.
\end{proof}


\noindent\textbf{Small-$n$ computation (evidence only).}
A direct search over $x$ and $y$ (bounding $y$ by $2/r$ once $r:=4/n-1/x$ is fixed) finds solutions for all $3\le n\le 1000$, e.g.
\[
\begin{array}{c|cccccccccc}
 n & 3&4&5&6&7&8&9&10&11&12\\\hline
(x,y,z) & (1,4,12)&(2,3,6)&(2,4,20)&(2,7,42)&(2,15,210)&(3,7,42)&(3,10,90)&(3,16,240)&(3,34,1122)&(4,13,156)
\end{array}
\]
This does not constitute a proof beyond $n\le 1000$.

\subsection*{VERIFICATION}
\begin{itemize}[leftmargin=2.2em]
\item Propositions 4--5 yield \emph{distinct} denominators and verify the exact algebra.
\item The scaling lemma preserves distinctness and ordering.
\item The computation is consistent with the known fact that any counterexample must be odd and not divisible by $3$ (i.e. $n\equiv 1,5\pmod 6$).
\end{itemize}

\subsection*{FINAL}
\textbf{UNRESOLVED.}

\begin{itemize}[leftmargin=2.2em]
\item[(i)] \textbf{Partial results obtained:}
  \begin{itemize}[leftmargin=2.2em]
  \item Gave explicit distinct Egyptian-fraction decompositions for all even $n>2$ and all $n>2$ divisible by $3$.
  \item Proved scaling and a reduction that any counterexample would force a prime counterexample.
  \item Verified computationally (by brute force) that solutions exist for all $3\le n\le 1000$.
  \end{itemize}
\item[(ii)] \textbf{Strongest obstruction / first gap:}
  The remaining unsolved case is $n$ coprime to $6$ (in particular primes $p\equiv 1,5\pmod 6$). No general parametrization covering all such $n$ is known.
\item[(iii)] \textbf{What would likely finish:}
  A complete set of modular identities covering all residue classes for primes, or a new structural argument producing a 3-term Egyptian decomposition for every $n$.
\item[(iv)] \textbf{Confidence assessment:}
  High confidence in the proved families; no claim about the full conjecture.
\end{itemize}

\subsection*{COMPLETION ESTIMATE}
35\%.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
