\section*{Erd\H{o}s Problem \#256}

\subsection*{1) FORMAL RESTATEMENT}
For each integer $n\ge 1$ and each nondecreasing $n$-tuple $(a_1,\dots,a_n)\in\mathbb{N}^n$ (i.e. $a_1\le\cdots\le a_n$), set
\[
P_{a_1,\dots,a_n}(z) \;:=\; \prod_{i=1}^n (1-z^{a_i}),\qquad z\in\mathbb{C}.
\]
For such a tuple define
\[
M(a_1,\dots,a_n)\;:=\;\max_{\left\lvert z\right\rvert=1}\,\left\lvert P_{a_1,\dots,a_n}(z)\right\rvert\in[0,\infty).
\]
(The maximum exists because $z\mapsto\left\lvert P_{a_1,\dots,a_n}(z)\right\rvert$ is continuous on the compact set $\{z\in\mathbb{C}:\left\lvert z\right\rvert=1\}$.)
Define
\[
f(n)\;:=\;\inf_{a_1\le\cdots\le a_n\in\mathbb{N}} M(a_1,\dots,a_n).
\]
Equivalently, $f(n)$ is the largest real number such that for \emph{every} choice of $a_1\le\cdots\le a_n\in\mathbb{N}$ one has
\[
\max_{\left\lvert z\right\rvert=1}\left\lvert \prod_{i=1}^n (1-z^{a_i})\right\rvert\;\ge\; f(n).
\]

The specific sub-question is the truth of the asymptotic lower bound
\[
(\star)\qquad \exists\,c>0\ \exists\,K>0\ \exists\,n_0\ \forall\,n\ge n_0:\ \log f(n)\ge K n^c.
\]
Here $\log$ denotes the natural logarithm; changing the base rescales $K$ by a constant factor.

\subsection*{2) QUICK LITERATURE/CONTEXT CHECK}
(From the standard summary on this problem and the cited papers.)
\begin{itemize}[leftmargin=2.2em]
\item Erd\H{o}s--Szekeres (1959) showed $f(n)>\sqrt{2n}$ and $\lim_{n\to\infty} f(n)^{1/n}=1$ \cite{ErSz59}.
\item A long sequence of upper bounds culminates in the result of Belov--Konyagin (1996):
\[\log f(n)\ll (\log n)^4\quad(n\to\infty),\]
which in particular gives a negative answer to $(\star)$ \cite{BeKo96}.
\item Very recently, Tang (2025) improved the classical general lower bound to
\[ f(n)\ge 2\sqrt{n}\quad\text{for all }n\in\mathbb{N},\]
the first improvement of the constant in the general lower bound since 1959 \cite{Tang25}.
\end{itemize}

\subsection*{3) ATTACK PLAN}
We target $(\star)$.  The plan is:
\begin{enumerate}[leftmargin=2.2em]
\item Use the known upper bound $\log f(n)\ll(\log n)^4$ (Belov--Konyagin).
\item Prove the elementary comparison $(\log n)^4 = o(n^c)$ for every fixed $c>0$.
\item Conclude that no fixed $c>0$ can satisfy $(\star)$.
\end{enumerate}

\subsection*{4) WORK}
\paragraph{Tiny cases / sanity checks.}
For $n=1$ and any $a_1\in\mathbb{N}$,
\[
M(a_1)=\max_{\left\lvert z\right\rvert=1}\left\lvert 1-z^{a_1}\right\rvert=2,
\]
because $\left\lvert 1-w\right\rvert\le 2$ for $\left\lvert w\right\rvert=1$ and equality holds at $w=-1$ (so take $z=e^{i\pi/a_1}$). Hence
\[
f(1)=2.
\]
More generally $f(n)\le 2^n$ by taking $a_1=\cdots=a_n=1$, which gives $(1-z)^n$ and $\max_{\left\lvert z\right\rvert=1}\left\lvert 1-z\right\rvert=2$.

As a numerical experiment (not a proof): restricting to $a_i\le 30$ and sampling $\theta$ on a fine grid for $z=e^{i\theta}$ suggests candidates giving
\[
M(4,8)\approx 3.0792\quad\text{(so }f(2)\le 3.0792\text{ in this restricted search)},
\]
\[
M(2,4,6)\approx 4.3899\quad\text{(so }f(3)\le 4.3899\text{ in this restricted search)}.
\]
These are only heuristic upper bounds for the true $f(n)$.

\paragraph{Known upper bound (input).}
We use the following result.
\begin{theorem}[Belov--Konyagin, consequence for $f(n)$ \cite{BeKo96}]
There exists an absolute constant $C>0$ such that for all $n\ge 2$,
\[
\log f(n)\le C\,(\log n)^4.
\]
Equivalently, $f(n)\le \exp\bigl(C(\log n)^4\bigr)$ for all $n\ge 2$.
\end{theorem}

\paragraph{Elementary comparison lemma.}
\begin{lemma}\label{lem:log4_vs_power}
For every $c>0$, one has $(\log n)^4 = o(n^c)$ as $n\to\infty$. Equivalently,
\[\lim_{n\to\infty}\frac{(\log n)^4}{n^c}=0.\]
\end{lemma}
\begin{proof}
Fix $c>0$ and set $\varepsilon:=c/8$. Consider the function $g(x)=x^{\varepsilon}/\log x$ for $x>e$.
Then
\[
\log g(x)=\varepsilon\log x-\log\log x\xrightarrow[x\to\infty]{}+\infty,
\]
so $g(x)\to\infty$. In particular, there exists $x_0$ such that for all $x\ge x_0$,
\[\log x \le x^{\varepsilon}.
\]
For $n\ge x_0$ this gives $(\log n)^4 \le n^{4\varepsilon}=n^{c/2}$, hence
\[
0\le \frac{(\log n)^4}{n^c}\le \frac{n^{c/2}}{n^c}=n^{-c/2}\xrightarrow[n\to\infty]{}0.
\]
\end{proof}

\paragraph{Disproof of $(\star)$.}
\begin{proposition}\label{prop:no_power_lower}
Statement $(\star)$ is false; i.e. there is \emph{no} constant $c>0$ such that $\log f(n)\gg n^c$.
\end{proposition}
\begin{proof}
Assume for contradiction that $(\star)$ holds. Then there exist $c>0$, $K>0$ and $n_0$ such that
$\log f(n)\ge K n^c$ for all $n\ge n_0$.

On the other hand, by the Belov--Konyagin bound there is $C>0$ such that for all $n\ge 2$,
$\log f(n)\le C(\log n)^4$.
By Lemma~\ref{lem:log4_vs_power}, $(\log n)^4=o(n^c)$, so we may choose $n_1$ such that for all $n\ge n_1$,
\[C(\log n)^4 < K n^c.
\]
Taking $n\ge \max\{n_0,n_1,2\}$ gives
\[\log f(n)\le C(\log n)^4 < K n^c \le \log f(n),
\]
a contradiction.
\end{proof}

\subsection*{5) VERIFICATION}
\begin{itemize}[leftmargin=2.2em]
\item The logical negation of $(\star)$ is: for every $c>0$ and every $K>0$ there exist arbitrarily large $n$ with $\log f(n)<K n^c$. Proposition~\ref{prop:no_power_lower} proves this, in fact giving $\log f(n)=O((\log n)^4)$ for all $n$.
\item Lemma~\ref{lem:log4_vs_power} is a standard calculus comparison and is proved explicitly.
\item The argument uses only: (i) the stated Belov--Konyagin upper bound, and (ii) the elementary fact that polylogarithms are $o(n^c)$.
\end{itemize}

\subsection*{6) FINAL}
\noindent LABEL: \textbf{FULL SOLUTION}\\
\noindent SUBLABEL: \textbf{COUNTEREXAMPLE/DISPROOF}\\
The claim ``$\exists c>0$ such that $\log f(n)\gg n^c$'' is false.  Indeed Belov--Konyagin prove
$\log f(n)\ll(\log n)^4$ \cite{BeKo96}, and since $(\log n)^4=o(n^c)$ for every $c>0$ (Lemma~\ref{lem:log4_vs_power}), no power lower bound $\log f(n)\gg n^c$ can hold.

\subsection*{7) COMPLETION ESTIMATE (MANDATORY)}
\noindent COMPLETION: 100\%.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
