\section*{Problem \#263}
\addcontentsline{toc}{section}{Problem \#263}

\subsection*{1. Formal Restatement}
Call a strictly increasing sequence of positive integers $(a_n)_{n\ge 1}$ a \emph{Type-2 irrationality sequence} if for every integer sequence $(b_n)_{n\ge 1}$ with
\[
\frac{b_n}{a_n}\to 1\qquad (n\to\infty),
\]
the series $\sum_{n\ge 1} 1/b_n$ is irrational (assuming the sum converges, which it does if $b_n\to\infty$ sufficiently fast).

Questions:
\begin{enumerate}[label=(\alph*),leftmargin=2em]
\item Is $a_n=2^{2^n}$ a Type-2 irrationality sequence?
\item Must every Type-2 irrationality sequence satisfy $a_n^{1/n}\to\infty$?
\end{enumerate}

\subsection*{2. Quick Literature / Context Check}
A ``folklore'' sufficient condition states that $\sum 1/a_n$ is irrational if $\lim a_n^{1/2^n}=\infty$.
Recent work of Kova\v{c} and Tao studies several ``types'' of irrationality sequences; in particular they show a broad class of sequences are \emph{not} Type-2 when $a_{n+1}/a_n^2\to 0$ (under convergence of $\sum 1/a_n$), while the folklore condition yields Type-2 examples under the much stronger growth $\liminf a_{n+1}/a_n^{2+\varepsilon}>0$.
The borderline case $a_{n+1}\asymp a_n^2$ includes $a_n=2^{2^n}$ (where $a_{n+1}=a_n^2$) and remains open.

\subsection*{3. Attack Plan}
Two natural directions:
\begin{enumerate}[leftmargin=2em]
\item \textbf{Try to prove Type-2 for $2^{2^n}$.} One would need a digit/carry or Diophantine approximation argument showing that any perturbation $b_n=2^{2^n}(1+o(1))$ forces the sum $\sum 1/b_n$ to avoid all rationals.
\item \textbf{Try to disprove Type-2 for $2^{2^n}$.} Construct a specific integer sequence $b_n$ with $b_n/2^{2^n}\to 1$ such that $\sum 1/b_n\in\mathbb{Q}$. This resembles building an ``Egyptian-fraction-like'' expansion with very large denominators constrained to lie in thin relative intervals.
\end{enumerate}

\subsection*{4. Work (provable partial results)}
We record the known partial implications as clean lemmas (with proofs only where elementary).

\begin{lemma}[A sufficient growth condition implying Type-2]
\label{lem:263-sufficient}
Assume there exists $\varepsilon>0$ and $c>0$ such that for all sufficiently large $n$,
\[
 a_{n+1} \ge c\, a_n^{2+\varepsilon}.
\]
Then $(a_n)$ is Type-2.
\end{lemma}

\begin{proof}[Proof sketch]
Under the hypothesis, $a_n$ grows faster than doubly-exponential in the sense that $a_n^{1/2^n}\to\infty$. If $b_n/a_n\to 1$ then also $b_n^{1/2^n}\to\infty$. The folklore irrationality criterion (see the cited sources) then gives that $\sum 1/b_n$ is irrational. Hence $(a_n)$ is Type-2.
\end{proof}

\begin{lemma}[A broad obstruction (Kova\v{c}--Tao)]
\label{lem:263-obstruction}
Suppose $(a_n)$ is strictly increasing, $\sum 1/a_n$ converges, and
\[
\frac{a_{n+1}}{a_n^2}\to 0.
\]
Then $(a_n)$ is \emph{not} Type-2.
\end{lemma}

\paragraph{Remark.}
The sequence $a_n=2^{2^n}$ satisfies $a_{n+1}/a_n^2=1$, so it lies exactly on the boundary not covered by Lemma~\ref{lem:263-obstruction} and does not satisfy the sufficient condition in Lemma~\ref{lem:263-sufficient}.

\subsection*{5. Verification / Sanity Checks}
\begin{itemize}[leftmargin=2em]
\item For $a_n=2^{2^n}$, indeed $a_{n+1}=2^{2^{n+1}}=(2^{2^n})^2=a_n^2$, so $a_{n+1}/a_n^2\equiv 1$.
\item The ``folklore'' criterion is consistent with classical irrationality proofs for rapidly growing Egyptian-fraction denominators.
\end{itemize}

\subsection*{6. FINAL}
\textbf{UNRESOLVED}

\subsubsection*{Why stuck / what remains open}
Existing theorems cover either much slower growth (showing \emph{not} Type-2) or much faster growth (showing Type-2). The sequence $2^{2^n}$ is precisely the ``critical'' growth rate $a_{n+1}\asymp a_n^2$ where current methods (per the cited references) do not decide Type-2.

The second question (whether Type-2 forces $a_n^{1/n}\to\infty$) also remains open; current results only give partial constraints.

\subsubsection*{Strongest partial result proved here}
Lemma~\ref{lem:263-sufficient}: a super-quadratic recurrence $a_{n+1}\ge c a_n^{2+\varepsilon}$ guarantees Type-2.

\subsubsection*{Next steps (concrete)}
\begin{enumerate}[leftmargin=2em]
\item Attempt a ``critical'' carry/approximation analysis for $a_{n+1}=a_n^2$ (or close): show that any rational approximation would force impossible congruence/size constraints on $b_n$.
\item Try to explicitly construct $b_n=2^{2^n}+o(2^{2^n})$ so that partial sums hit a fixed rational within an error smaller than the remaining tail.
\item Search for intermediate criteria between $a_{n+1}/a_n^2\to 0$ and $a_{n+1}\ge a_n^{2+\varepsilon}$ that can decide the borderline.
\end{enumerate}

\subsection*{7. Completion Estimate}
\textbf{35\%}. We can cleanly state the known sufficient/obstruction regimes and locate $2^{2^n}$ at the critical boundary, but the main question is still open.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{thebibliography}{99}

\bibitem{Bloom263}
T. F. Bloom, \emph{Erd\H{o}s Problem \#263}, \url{https://www.erdosproblems.com/263} (accessed 2026-01-17).

\bibitem{BoLo90}
P. Borwein and L. Loring,
\emph{Some questions of Erd\H{o}s and Graham on numbers of the form $\sum g_n/2^{g_n}$},
Amer. Math. Monthly \textbf{97} (1990), 22--26.

\bibitem{TUZ20}
S. Tengely, M. Ulas, and J. Zygad\l{}o,
\emph{On a Diophantine equation of Erd\H{o}s and Graham},
Revista Colombiana de Matem\'aticas \textbf{54} (2020), 445--459.

\bibitem{KoTa24}
V. Kova\v{c} and T. Tao,
\emph{The Ahmes series and irrationality sequences},
arXiv:2406.17593 (version v4, 2025-07-14).

\end{thebibliography}

