FORMAL RESTATEMENT
A set $S\subseteq\{1,\dots,N\}$ is a \emph{Sidon set} if all sums $s_i+s_j$ with $s_i\le s_j$ are distinct.
Let $h(N)$ denote the maximum possible size of a Sidon set in $\{1,\dots,N\}$.
The question asks whether for every $\epsilon>0$ one has
\[
 h(N)=N^{1/2}+O_{\epsilon}(N^{\epsilon}).
\]

QUICK LITERATURE/CONTEXT CHECK
The problem text records the classical upper bound $h(N)\le N^{1/2}+N^{1/4}+1$ (Erdős--Turán, Lindström) and a sequence of small improvements to the $N^{1/4}$ coefficient, as well as record computations.
No further resolution is given in the problem text, so I treat the $O_{\epsilon}(N^{\epsilon})$ error term as open here.

ATTACK PLAN
The natural approaches are:
(1) improve the classical upper bound toward $\sqrt{N}+o(N^{\delta})$ via more efficient counting of sums/differences, possibly using Fourier-analytic or incidence-geometric methods;
(2) build explicit constructions with $|S|\ge \sqrt{N}-o(\sqrt{N})$ to show tightness and control the second-order term.
A disproof would require exhibiting infinitely many $N$ where $h(N)\ge \sqrt{N}+N^{c}$ for some fixed $c>0$.

WORK
\textit{Fast reality check (exact search for small $N$).}
I ran a backtracking search maximizing $|S|$ subject to distinct positive differences.
The exact values found for $1\le N\le 40$ are:
\[
\begin{array}{c|cccccccccc}
N&1&2&3&4&5&6&7&8&9&10\\\hline
h(N)&1&2&2&3&3&3&4&4&4&4
\end{array}
\qquad
\begin{array}{c|cccccccccc}
N&11&12&13&14&15&16&17&18&19&20\\\hline
h(N)&4&5&5&5&5&5&5&6&6&6
\end{array}
\]
\[
\begin{array}{c|cccccccccc}
N&21&22&23&24&25&26&27&28&29&30\\\hline
h(N)&6&6&6&6&6&7&7&7&7&7
\end{array}
\qquad
\begin{array}{c|cccccccccc}
N&31&32&33&34&35&36&37&38&39&40\\\hline
h(N)&7&7&7&7&8&8&8&8&8&8
\end{array}
\]

\medskip
Lemma 30.1 (elementary upper bound via differences).
If $S\subseteq\{1,\dots,N\}$ is Sidon and $|S|=m$, then
\[
\binom{m}{2}\le N-1\qquad\text{and hence}\qquad m\le \frac{1+\sqrt{8N-7}}{2}.
\]

Proof.
Write $S=\{s_1<\dots<s_m\}$.
Sidon implies that all positive differences $s_j-s_i$ with $1\le i<j\le m$ are distinct: indeed, if
$s_j-s_i=s_{j'}-s_{i'}$ with $(i,j)\ne(i',j')$, then rearranging gives the repeated sum
$s_j+s_{i'}=s_{j'}+s_i$.
There are $\binom{m}{2}$ positive differences, and each lies in $\{1,2,\dots,N-1\}$, a set of size $N-1$.
Therefore $\binom{m}{2}\le N-1$.
Solving $m(m-1)/2\le N-1$ yields $m\le (1+\sqrt{8N-7})/2$.
\qed

\medskip
Lemma 30.2 (parabola is Sidon in $\mathbb{Z}_p\times\mathbb{Z}_p$).
Let $p$ be prime and let
\[
\mathcal{S}:=\{(t,t^2): t\in\mathbb{Z}_p\}\subseteq \mathbb{Z}_p\times\mathbb{Z}_p.
\]
Then $\mathcal{S}$ is Sidon in the sense that if
\((t_1,t_1^2)-(t_2,t_2^2)=(t_3,t_3^2)-(t_4,t_4^2)\)
holds in $\mathbb{Z}_p\times\mathbb{Z}_p$, then $t_1=t_3$ and $t_2=t_4$.

Proof.
The equality of differences gives the system in $\mathbb{Z}_p$:
\[
 t_1-t_2=t_3-t_4=:d,\qquad t_1^2-t_2^2=t_3^2-t_4^2.
\]
If $d=0$, then $t_1=t_2$ and $t_3=t_4$, so the pairs are equal.
Assume $d\ne 0$. Factor the second equality:
\[
(t_1-t_2)(t_1+t_2)=(t_3-t_4)(t_3+t_4).
\]
Since $t_1-t_2=t_3-t_4=d\ne 0$ in the field $\mathbb{Z}_p$, we may cancel $d$ to obtain $t_1+t_2=t_3+t_4$.
Now solve the linear system
$t_1-t_2=t_3-t_4$ and $t_1+t_2=t_3+t_4$ in $\mathbb{Z}_p$:
adding gives $2t_1=2t_3$, hence $t_1=t_3$, and subtracting gives $2t_2=2t_4$, hence $t_2=t_4$.
\qed

\medskip
Lemma 30.3 (integer Sidon construction of size $p$ in $[0,p^2-1]$).
Let $p$ be prime and define
\[
S_p:=\{\,t + p\,(t^2\bmod p):\ t=0,1,\dots,p-1\,\}\subseteq\{0,1,\dots,p^2-1\}.
\]
Then $S_p$ is a Sidon set in $\{0,1,\dots,p^2-1\}$, and in particular $h(p^2)\ge p$.

Proof.
Write $f(t)=t+p\,(t^2\bmod p)$.
Suppose $f(t_1)-f(t_2)=f(t_3)-f(t_4)$ as integers.
Then
\[
(t_1-t_2)+p\big((t_1^2\bmod p)-(t_2^2\bmod p)\big)
=
(t_3-t_4)+p\big((t_3^2\bmod p)-(t_4^2\bmod p)\big).
\]
Reducing modulo $p$ gives $t_1-t_2\equiv t_3-t_4\pmod p$.
But each difference $t_i-t_j$ lies in $\{-(p-1),\dots,p-1\}$, so congruence modulo $p$ forces equality in $\mathbb{Z}$:
$t_1-t_2=t_3-t_4$.
Subtracting this common integer from both sides and dividing by $p$ yields
\((t_1^2\bmod p)-(t_2^2\bmod p)=(t_3^2\bmod p)-(t_4^2\bmod p)\) in $\mathbb{Z}$, hence also modulo $p$.
Therefore, in $\mathbb{Z}_p\times\mathbb{Z}_p$ we have
\[
(t_1, t_1^2) - (t_2, t_2^2) = (t_3, t_3^2) - (t_4, t_4^2).
\]
By Lemma 30.2, this implies $t_1=t_3$ and $t_2=t_4$.
Thus all positive differences in $S_p$ are distinct, so $S_p$ is Sidon.
\qed

VERIFICATION
Lemma 30.1 uses only the equivalence between repeated positive differences and repeated sums, and a counting bound.
Lemma 30.2 is a direct field calculation.
Lemma 30.3 carefully uses the ``no carry'' uniqueness coming from representing integers in base $p$ with coefficients in $[0,p-1]$.
The small-$N$ values of $h(N)$ were computed by exhaustive backtracking and are internally consistent with the trivial upper bound from Lemma 30.1.

FINAL
**UNRESOLVED**
(i) Strongest proved partial result here (from the WORK): explicit constructions give $h(p^2)\ge p$ for prime $p$ (Lemma 30.3), i.e. $h(N)\ge (1+o(1))\sqrt{N}$ along squares of primes, while an elementary argument gives $h(N)\le (\sqrt{2}+o(1))\sqrt{N}$ (Lemma 30.1). The conjectured refinement $h(N)=\sqrt{N}+O_{\epsilon}(N^{\epsilon})$ is far beyond these elementary bounds.
(ii) First gap: improve the upper bound to $h(N)\le \sqrt{N}+o(N^{\delta})$ for every fixed $\delta>0$ (or even to $\sqrt{N}+O(N^{1/4})$ by elementary means). The stated conjecture asks for error $O_{\epsilon}(N^{\epsilon})$.
(iii) Top 3 next moves:
  (1) Reproduce (and then try to sharpen) the classical Erdős--Turán/Lindström argument giving $\sqrt{N}+O(N^{1/4})$, tracking precisely where the $N^{1/4}$ term appears.
  (2) Search computationally for extremal Sidon sets at moderate $N$ and fit $h(N)-\sqrt{N}$ to guess the true second-order term.
  (3) Translate the Sidon condition to an incidence geometry statement (distinct sums/differences) and seek sharper incidence bounds tailored to the grid $\{1,\dots,N\}$.
(iv) Minimal counterexample structure: if the conjecture fails, then there exists $\epsilon_0>0$ and infinitely many $N$ with $h(N)\ge \sqrt{N}+cN^{\epsilon_0}$; such extremal sets would need unusually high ``packing efficiency'' among differences, suggesting additional algebraic structure beyond known constructions.


