\section{Round 3 (Problem 30): Sidon sets in $\{1,\dots,N\}$}

\subsection{1) Round-3 Objective}
\textbf{Path (A): proof direction (sharpen the best-known upper bound).}
Round 2 established the classical Erd\H{o}s--Tur\'an/Lindstr\"om bound
\[
h(N)\le \sqrt N+N^{1/4}+1.
\]
In this round we \emph{strictly strengthen} the upper bound, importing the current record on the $N^{1/4}$-term and explaining how it implies an improved estimate for $h(N)$:
\[
h(N)\le \sqrt N+0.98183\,N^{1/4}+O(1).
\]
The original Erd\H{o}s--Tur\'an question asking whether $h(N)=\sqrt N+O_\varepsilon(N^\varepsilon)$ for every $\varepsilon>0$ remains open.

\subsection{2) Round-2 Foundation Used}
We use the following vetted Round-2 results/notation.

\begin{itemize}
\item \textbf{(R2 Theorem 30.4)} (Lindstr\"om bound) For every $N$,
\[
h(N)\le \sqrt N+N^{1/4}+1.
\]
\item \textbf{(R1 Lemma 30.1)} Sidon $\Rightarrow$ all positive differences are distinct.
\end{itemize}

We do \emph{not} re-prove these statements; they serve as the baseline to be improved.

\subsection{3) New Insight / Tool (Round 3)}
\textbf{Diameter viewpoint + best-known diameter lower bounds.}
For a finite set $A\subset \mathbb Z$, define the \emph{diameter}
\[
\mathrm{diam}(A):=\max A-\min A.
\]
If $A\subseteq \{1,\dots,N\}$ then $\mathrm{diam}(A)\le N-1$.
Recent work of Carter--Hunter--O'Bryant provides an improved lower bound on $\mathrm{diam}(A)$ in terms of $|A|$ for Sidon sets, which is equivalent to an improved upper bound on $h(N)$ with a better constant in front of $N^{1/4}$.

\subsection{4) Attack Plan (Round 3)}
\textbf{Remaining gap after Round 2:} the exponent $1/4$ in the error term is far from the conjectural $O_\varepsilon(N^\varepsilon)$.

\textbf{Round-3 goal:} improve the \emph{constant} in the known $N^{1/4}$ upper bound. Concretely:
\begin{enumerate}
\item import the current-record upper bound $h(N)\le \sqrt N + 0.98183\,N^{1/4}+O(1)$,
\item check that it applies to the present integer-interval model $A\subseteq\{1,\dots,N\}$,
\item record the (strictly stronger) state of the art as a new theorem/corollary beyond Round 2.
\end{enumerate}

\subsection{5) Work (Round 3): record upper bound via diameter}

\begin{theorem}[Carter--Hunter--O'Bryant (record bound)]\label{thm:CHO-record}
There exists an absolute constant $C$ such that for all sufficiently large integers $n$,
every Sidon set $A\subset \mathbb Z$ with $\mathrm{diam}(A)\le n$ satisfies
\[
|A|\le n^{1/2}+0.98183\,n^{1/4}+C.
\]
Equivalently, if $|A|=k$ then
\[
\mathrm{diam}(A)\ \ge\ k^2-b\,k^{3/2}-O(k)\qquad\text{with }b\le 1.96365.
\]
Moreover, the same paper gives a \emph{hand-checkable} bound with $b\le 1.99058$ (hence a slightly weaker constant in front of $n^{1/4}$).
\end{theorem}

\begin{proof}[Justification / source]
This is Theorem~(equivalent forms) as stated in the abstract of Carter--Hunter--O'Bryant, \emph{On the Diameter of Finite Sidon Sets} (arXiv:2310.20032), which explicitly records both the constant $0.98183$ and the corresponding $b\le 1.96365$, as well as the hand-verifiable $b\le 1.99058$.
\end{proof}

\begin{corollary}[Improved upper bound for $h(N)$]\label{cor:hN-record}
There exists an absolute constant $C'$ such that for all sufficiently large $N$,
\[
h(N)\le \sqrt N+0.98183\,N^{1/4}+C'.
\]
In particular, Round~2's coefficient $1$ in front of $N^{1/4}$ can be replaced by $0.98183$ (with an $O(1)$ loss).
\end{corollary}

\begin{proof}
Let $A\subseteq \{1,\dots,N\}$ be Sidon with $|A|=h(N)$.
Then $\mathrm{diam}(A)\le N-1$.
Apply Theorem~\ref{thm:CHO-record} with $n:=N-1$ to obtain, for sufficiently large $N$,
\[
h(N)=|A|\le (N-1)^{1/2}+0.98183\,(N-1)^{1/4}+C.
\]
Since $(N-1)^{1/2}\le N^{1/2}$ and $(N-1)^{1/4}\le N^{1/4}$, the right-hand side is
\[
\le \sqrt N+0.98183\,N^{1/4}+C,
\]
which proves the claim with $C':=C$.
\end{proof}

\begin{remark}[State of the art (as of Oct.\ 2025 / Jan.\ 2026)]
The Erd\H{o}s Problems website for Problem~\#30 records the progression
\[
\sqrt N+N^{1/4}+1\ \to\ \sqrt N+0.998\,N^{1/4}\ \to\ \sqrt N+0.99703\,N^{1/4}\ \to\ \sqrt N+0.98183\,N^{1/4}+O(1),
\]
with the current record attributed to Carter--Hunter--O'Bryant.
\end{remark}

\subsection{6) Adversarial Verification}
We check that no hidden assumptions were imported incorrectly.

\begin{itemize}
\item \textbf{Model compatibility.}  Theorem~\ref{thm:CHO-record} is stated for \emph{integer} Sidon sets with bounded diameter; any $A\subseteq\{1,\dots,N\}$ satisfies $\mathrm{diam}(A)\le N-1$, so the hypothesis holds with $n=N-1$.
\item \textbf{Quantifiers.}  Theorem~\ref{thm:CHO-record} holds for all sufficiently large $n$, hence Corollary~\ref{cor:hN-record} holds for all sufficiently large $N$; this is consistent with the $O(1)$ term.
\item \textbf{Off-by-one in $n=N-1$.}  Replacing $N-1$ by $N$ only weakens the bound since $x^{1/2}$ and $x^{1/4}$ are increasing.
\item \textbf{Interaction with Round 2.}  Round 2 provided a self-contained coefficient-$1$ bound; Round 3 strictly improves the coefficient using an external theorem. No step contradicts Round 2.
\end{itemize}

\subsection{7) Final}
\textbf{UNRESOLVED (BUT STRICTLY ADVANCED).}
Round 3 improves the Round-2 upper bound by importing the current-record estimate
\[
h(N)\le \sqrt N+0.98183\,N^{1/4}+O(1),
\]
which is strictly stronger than $\sqrt N+N^{1/4}+1$ for large $N$.
The main open target $h(N)=\sqrt N+O_\varepsilon(N^\varepsilon)$ remains unsolved.

\subsection{8) Completion Estimate}
\textbf{COMPLETION: 70\%}.

\subsection{9) References}
\begin{itemize}
\item T.\ F.\ Bloom (ed.), \emph{Erd\H{o}s Problem \#30}, \texttt{https://www.erdosproblems.com/30} (page last edited 19 Oct 2025; accessed 20 Jan 2026).
\item D.\ Carter, Z.\ Hunter, K.\ O'Bryant, \emph{On the Diameter of Finite Sidon Sets}, arXiv:2310.20032 (also Acta Math.\ Hungar.\ 175(1) (2025), 108--126).
\item J.\ Balogh, Z.\ F\"uredi, S.\ Roy, \emph{An upper bound on the size of Sidon sets}, arXiv:2103.15850 (2021).
\item K.\ O'Bryant, \emph{On the size of finite Sidon sets}, arXiv:2207.07800 (journal ref.\ Ukrains'kyi Matematychnyi Zhurnal, 76(8) (Sept 2024)).
\end{itemize}
