% Erdos Problem #470
% URL: https://www.erdosproblems.com/470

\noindent\textbf{1) FORMAL RESTATEMENT.}

Let $\sigma(n)$ be the sum of positive divisors of $n$. Call $n$ \emph{weird} if
\[
\sigma(n)\ge 2n
\]
(i.e. $n$ is abundant or perfect) and $n$ is \emph{not} pseudoperfect, i.e. $n$ is not the sum of any set of \emph{distinct} proper divisors of $n$.

Call $n$ \emph{primitive weird} if $n$ is weird and no proper divisor of $n$ is weird.

Questions:
\begin{enumerate}
\item Do there exist any odd weird numbers?
\item Are there infinitely many primitive weird numbers?
\end{enumerate}

\medskip
\noindent\textbf{2) QUICK LITERATURE/CONTEXT CHECK.}

The problem text states several known results: Benkoski--Erd\H{o}s \cite{BeEr74} proved weird numbers have positive density; Melfi \cite{Me15} proved conditional infinitude of primitive weird numbers under a prime-gap hypothesis; Fang \cite{Fa22} showed no odd weird numbers below $10^{21}$; Liddy--Riedl \cite{LiRi18} proved an odd weird number must have at least 6 prime divisors. I will not assume any of these beyond what is explicitly stated.

\medskip
\noindent\textbf{3) ATTACK PLAN.}

\begin{itemize}
\item Give a complete verification that $70$ is weird (and primitive weird) directly from its divisor set.
\item Run an explicit search for weird numbers in a moderate range to sanity-check parity and primitiveness phenomena.
\item Isolate crisp structural constraints that any odd weird number would have to satisfy (based only on definitions and verified computations).
\end{itemize}

\medskip
\noindent\textbf{4) WORK.}

\noindent\textbf{Lemma 470.1 ($70$ is weird and primitive weird).}
The integer $70$ is weird, and no proper divisor of $70$ is weird.

\textit{Proof.}
The positive divisors of $70$ are
\[
1,2,5,7,10,14,35,70.
\]
Hence
\[
\sigma(70)=1+2+5+7+10+14+35+70=144\ge 140=2\cdot 70,
\]
so $70$ is abundant.

To check that $70$ is \emph{not} pseudoperfect, consider its proper divisors
\[
1,2,5,7,10,14,35.
\]
If a subset of these sums to $70$, it must include $35$ (since the sum of the six proper divisors excluding $35$ is $1+2+5+7+10+14=39<70$). But if $35$ is included, the remaining needed sum is $35$. The largest possible sum using $\{1,2,5,7,10,14\}$ is $39$, and a direct check of combinations shows no subset sums to $35$:
\[
14+10+7+2+1=34,\quad 14+10+7+5=36,
\]
and all other subsets are smaller than $34$ or exceed $36$. Hence $35$ cannot be completed to $70$, so no subset of proper divisors sums to $70$. Therefore $70$ is weird.

Finally, every proper divisor of $70$ is one of $1,2,5,7,10,14,35$. Each has $\sigma(m)<2m$ (so is not abundant), hence cannot be weird. Thus $70$ is primitive weird. \hfill$\square$

\medskip
\noindent\textbf{Lemma 470.2 (computational parity check up to $200000$).}
There are no odd weird numbers $\le 200000$. Moreover, in this range there are $395$ weird numbers and $14$ primitive weird numbers.

\textit{Proof (explicit computation).}
For each $2\le n\le 200000$ I computed the list of proper divisors, checked abundance by comparing $\sum_{d\mid n,\ d<n} d$ with $n$, and checked pseudoperfectness by exact subset-sum DP over proper divisors (bitset method). A number was declared weird iff it was abundant and not pseudoperfect; it was declared primitive weird iff it was weird and no proper divisor was weird.

The results were:
\begin{verbatim}
count weird = 395
count primitive weird = 14
first 15 weird:
[70, 836, 4030, 5830, 7192, 7912, 9272, 10430, 10570, 10792,
 10990, 11410, 11690, 12110, 12530]
first 14 primitive weird:
[70, 836, 4030, 5830, 7192, 7912, 9272, 10792, 17272, 45356,
 73616, 83312, 91388, 113072]
odd weird found: []
\end{verbatim}
So no odd weird numbers occur up to $200000$. \hfill$\square$

\medskip
\noindent\textbf{5) VERIFICATION.}

\begin{itemize}
\item Lemma~470.1 is a complete hand-check using the explicit divisor list of $70$.
\item Lemma~470.2 is an exhaustive check up to $200000$ using exact arithmetic and exact subset-sum DP, so it is a rigorous finite computation.
\item The computational evidence is consistent with the statement in the problem text that no odd weird numbers exist below $10^{21}$ (though I did not verify that large bound).
\end{itemize}

\medskip
\noindent\textbf{6) FINAL.} \textbf{UNRESOLVED}

(i) \textbf{Strongest proved partial result.}
We have a complete proof that $70$ is (primitive) weird (Lemma~470.1), and an exhaustive search showing there are no odd weird numbers $\le 200000$ and listing the first several weird and primitive weird numbers (Lemma~470.2). The problem statement further records deep results (positive density, conditional infinitude of primitive weirds, and large-scale nonexistence results for odd weirds), but these are not proved here.

(ii) \textbf{First gap (crisp).}
Either produce an explicit odd weird number (counterexample to ``none exist''), or prove that no odd weird numbers exist. Independently, prove or disprove that there are infinitely many primitive weird numbers without additional hypotheses.

(iii) \textbf{Top 3 next moves.}
\begin{enumerate}
\item Try to find an unconditional infinite family of primitive weird numbers by proving a sufficient condition for weirdness in a parametrized family (e.g. numbers with controlled divisor structure), and then verifying primitiveness.
\item If aiming to rule out odd weirds, develop constraints for an odd abundant number to be non-pseudoperfect (subset-sum obstruction) and attempt to show these constraints cannot hold.
\item Extend computation: push the ``no odd weird'' search to higher bounds and analyze the prime-factor patterns of odd abundant numbers that fail/passed pseudoperfectness tests, to guide a theoretical obstruction.
\end{enumerate}

(iv) \textbf{Minimal counterexample structure.}
A minimal odd weird number (if it exists) would be an odd abundant number whose proper divisors fail every subset-sum representation of $n$; by the cited result in the problem text it would have at least 6 prime divisors. Any attempt to prove nonexistence must explain why every odd abundant integer with sufficiently many prime factors is nevertheless pseudoperfect.
