%Erdos problem 517
\noindent\textbf{FORMAL RESTATEMENT}

Let $(n_k)_{k\ge 1}$ be a strictly increasing sequence of nonnegative integers and let $(a_k)_{k\ge 1}$ be nonzero complex numbers such that the power series
\[
f(z)=\sum_{k=1}^{\infty} a_k z^{n_k}
\]
converges for all $z\in\mathbb{C}$ (so $f$ is entire). Assume the gap condition
\[
\frac{n_k}{k}\to\infty\quad\text{as }k\to\infty.
\]
Question: Must it follow that $f$ assumes every complex value infinitely often (i.e., for every $w\in\mathbb{C}$, the equation $f(z)=w$ has infinitely many solutions in $\mathbb{C}$)?

Edge cases: $f$ is transcendental because infinitely many terms are present. By classical Picard theory, there is at most one value $w$ that can fail to be assumed infinitely often; the conjecture asks that there is no such exceptional value.

\medskip
\noindent\textbf{QUICK LITERATURE/CONTEXT CHECK}

The problem statement records: Fej\'er proved that if $\sum 1/n_k<\infty$ then $f$ assumes every value at least once; Biernacki proved under the same hypothesis every value is assumed infinitely often. P\'olya proved that if $f$ has finite order then, assuming $\limsup (n_{k+1}-n_k)=\infty$, $f$ assumes every value infinitely often.

\medskip
\noindent\textbf{ATTACK PLAN}

\begin{itemize}
\item Proof-track idea: show that the gap condition prevents $f$ from having an exceptional value $w$ with only finitely many preimages, by ruling out representations $f-w = P e^g$.
\item Disproof-track idea: attempt to explicitly construct a lacunary entire function with $n_k/k\to\infty$ that omits a value or has only finitely many preimages for some value (i.e., build $f-w=P e^g$ with lacunary Maclaurin support).
\end{itemize}
I did not complete either track.

\medskip
\noindent\textbf{WORK}

\noindent\textbf{Fast reality check (the gap condition forces unbounded successive gaps).}
A quick check shows $n_k/k\to\infty$ implies $n_{k+1}-n_k$ cannot be bounded.

\noindent\textbf{Additional computational sanity check (toy lacunary series).}
Consider the lacunary entire function $f(z)=\sum_{k\ge 1} \frac{z^{2^k}}{(2^k)!}$ (here $n_k=2^k$, so $n_k/k\to\infty$). Truncating at $k=4$ gives
\[
P(z)=\frac{z^2}{2!}+\frac{z^4}{4!}+\frac{z^8}{8!}+\frac{z^{16}}{16!}.
\]
Using \texttt{numpy.roots} on the degree-$16$ coefficient array, I obtained:
\begin{itemize}
\item degree $16$; $16$ roots in total; $2$ roots at $0$ (numerically $|z|<10^{-8}$, consistent with the factor $z^2$);
\item smallest nonzero root modulus $\approx 3.3422052103$ (approximately at $\pm 3.3422052103\, i$);
\item largest root modulus $\approx 12.2987828606$.
\end{itemize}
This does not resolve the conjecture; it only checks that a representative lacunary pattern under the hypothesis behaves nontrivially in small truncations.


\medskip
\noindent\textbf{Lemma 1 (unbounded successive gaps).}
If $n_k/k\to\infty$ as $k\to\infty$, then
\[
\limsup_{k\to\infty} (n_{k+1}-n_k)=\infty.
\]

\textit{Proof.}
Suppose for contradiction that $n_{k+1}-n_k\le D$ for all $k$ for some fixed $D$. Then by telescoping,
\[
n_k \le n_1 + D(k-1) \le (n_1+D)k.
\]
Dividing by $k$ gives $n_k/k\le n_1+D$, contradicting $n_k/k\to\infty$.
\hfill$\square$

\medskip
\noindent\textbf{Lemma 2 (finite preimages imply an exponential-polynomial factorisation).}
Let $F$ be a nonconstant entire function and $w\in\mathbb{C}$. If the equation $F(z)=w$ has only finitely many solutions in $\mathbb{C}$ (counted without multiplicity), then there exist a nonzero polynomial $P$ and an entire function $g$ such that
\[
F(z)-w = P(z) e^{g(z)}.
\]

\textit{Proof.}
Let $G(z):=F(z)-w$. By assumption, $G$ is an entire function with only finitely many zeros (counted with multiplicity). Let those zeros be $\zeta_1,\dots,\zeta_m$ with multiplicities $m_1,\dots,m_m$.
Define the polynomial
\[
P(z):=\prod_{j=1}^m (z-\zeta_j)^{m_j}.
\]
Then $H(z):=G(z)/P(z)$ is entire and has no zeros: all zeros of $G$ have been divided out with the correct multiplicity. Since $H$ is holomorphic and nonvanishing on the simply connected domain $\mathbb{C}$, it admits an entire logarithm: define
\[
 g(z):=\int_0^z \frac{H'(\zeta)}{H(\zeta)}\,d\zeta.
\]
Because $H'/H$ is entire (holomorphic with no poles) and $\mathbb{C}$ is simply connected, the integral is path-independent and yields an entire function $g$ with $g'(z)=H'(z)/H(z)$. Then $(e^{-g}H)'=0$ (differentiate to check), so $H(z)=C e^{g(z)}$ for some constant $C\neq 0$. Absorbing $C$ into $P$ (replace $P$ by $CP$) gives $G=P e^{g}$.
\hfill$\square$

\medskip
\noindent\textbf{VERIFICATION}

\begin{itemize}
\item Lemma 2 is a standard factorisation for entire functions with finitely many zeros; I verified the logarithm construction carefully: $H$ nonvanishing on simply connected $\mathbb{C}$ implies $H'/H$ entire and integrable path-independently.
\item Neither lemma resolves whether such a factorisation can be compatible with lacunary Maclaurin support $\{n_k\}$ and the condition $n_k/k\to\infty$.
\end{itemize}

\medskip
\noindent\textbf{FINAL}

\textbf{UNRESOLVED}

(i) Strongest proved partial result: The gap condition $n_k/k\to\infty$ forces $\limsup (n_{k+1}-n_k)=\infty$ (Lemma 1). If $f$ had a value $w$ that is assumed only finitely often, then $f-w$ must factor as $P e^{g}$ with $P$ a polynomial and $g$ entire (Lemma 2).

(ii) First gap (crisp): Show (or refute) that an entire lacunary series with $n_k/k\to\infty$ cannot admit a decomposition $f-w=P e^{g}$ for any $w\in\mathbb{C}$, where $P$ is a polynomial and $g$ is entire.

(iii) Top 3 next moves:
\begin{enumerate}
\item Coefficient-obstruction attempt: derive necessary algebraic/differential recurrences for Maclaurin coefficients of $P e^{g}$ and prove these recurrences force the support to have $n_k=O(k)$ (or at least violate $n_k/k\to\infty$).
\item Computational search for structured candidates: attempt to explicitly construct (via solving for coefficients) a nonvanishing entire function with sparse support $\{n_k\}$, starting from a desired lacunary series and taking a convergent logarithm $g=\log f$.
\item Use growth/order: combine Lemma 1 (unbounded gaps) with value-distribution theorems that rule out exceptional values for functions with sufficiently sparse exponents unless the order is finite (where P\'olya's result applies).
\end{enumerate}

(iv) Minimal counterexample structure (if the conjecture is false): would be an entire $f(z)=\sum a_k z^{n_k}$ with $n_k/k\to\infty$ for which some value $w$ has only finitely many preimages, hence $f-w=P e^{g}$; such a counterexample would amount to a nonvanishing entire function (up to a polynomial factor) whose Maclaurin series has Fabry gaps.

