% Erdos Problem #602
% URL: https://www.erdosproblems.com/602

Let $(A_i)$ be a family of sets with $\lvert A_i\rvert=\aleph_0$ for all $i$, such that for any $i\neq j$ we have $\lvert A_i\cap A_j\rvert$ finite and $\neq 1$. Is there a $2$-colouring of $\cup A_i$ such that no $A_i$ is monochromatic? A problem of Komj\'{a}th. The existence of such a $2$-colouring is sometimes known as Property B.

\noindent\textbf{1) FORMAL RESTATEMENT}\par
Let $I$ be an index set and let $(A_i)_{i\in I}$ be a family of sets with $|A_i|=\aleph_0$ for all $i$.
Assume that for all distinct $i,j\in I$, the intersection $A_i\cap A_j$ is finite and has size not equal to $1$.

Question: does there exist a function (2-colouring)
\[
 c: \bigcup_{i\in I} A_i \to \{0,1\}
\]
such that for every $i\in I$, the restriction $c|_{A_i}$ is not constant (equivalently, $A_i$ contains at least one point of each colour)?

\noindent\textbf{2) QUICK LITERATURE/CONTEXT CHECK}\par
The statement notes this is known as Property B in this setting. No further results are stated, and I do not import external literature.

\noindent\textbf{3) ATTACK PLAN}\par
\begin{itemize}
\item Proof-track idea: attempt a transfinite greedy colouring, using the fact that each $A_i$ is infinite and pairwise intersections are finite. The main difficulty is that at an uncountable stage, $A_i$ might already have many coloured points.
\item Disproof-track idea: search for an uncountable family with the given intersection pattern that forces one $A_i$ to become monochromatic under any 2-colouring.
\item Sanity check: solve the countable-indexed case explicitly.
\end{itemize}

\noindent\textbf{4) WORK}\par
\textbf{Fast reality check (countable family).} If $I$ is countable (in particular finite), a direct greedy construction works and does not even need the intersection hypotheses.

\medskip
\noindent\textbf{Lemma 602.1 (countable family of countably infinite sets has Property B).}\\
Assume $I$ is countable. Then there exists a 2-colouring of $\bigcup_{i\in I}A_i$ such that no $A_i$ is monochromatic.

\noindent\emph{Proof.}
Enumerate $I$ as $\{i_0,i_1,i_2,\dots\}$.
We build a colouring $c$ in stages. Maintain a finite set $S_n$ of already coloured points after stage $n$, with $S_0=\emptyset$.

At stage $n$, since $A_{i_n}$ is countably infinite and $S_n$ is finite, choose two distinct points
\[
x_n, y_n \in A_{i_n}\setminus S_n.
\]
Define $c(x_n)=0$ and $c(y_n)=1$, and leave all other as-yet-uncoloured points undefined for now.
Set $S_{n+1}:=S_n\cup\{x_n,y_n\}$.

After completing all stages, define $c$ on any remaining uncoloured points arbitrarily (e.g., colour them $0$).

Now fix any $i=i_n$. By construction, $x_n,y_n\in A_i$ and $c(x_n)=0$, $c(y_n)=1$. Hence $A_i$ contains both colours and is not monochromatic.
This holds for every $i\in I$, so $c$ is a valid 2-colouring. \qed

\medskip
\noindent\textbf{Lemma 602.2 (transfinite recursion under ``infinitely many new points'').}\\
Assume $I$ is well-ordered and that for every $i\in I$,
\[
A_i\setminus\bigcup_{j<i}A_j\text{ is infinite.}
\]
Then there exists a 2-colouring of $\bigcup_{i\in I}A_i$ such that no $A_i$ is monochromatic.

\noindent\emph{Proof.}
We define a colouring by transfinite recursion along the well-order of $I$.
Suppose colours have been assigned to all points in $\bigcup_{j<i}A_j$.
By hypothesis, the set of ``new'' points
\[
B_i:=A_i\setminus\bigcup_{j<i}A_j
\]
is infinite, hence contains two distinct uncoloured points $x_i,y_i$.
Assign $c(x_i)=0$ and $c(y_i)=1$.
Leave any other points of $B_i$ either uncoloured until the end or colour them arbitrarily.

At the end, colour any remaining uncoloured points arbitrarily.

For each $i$, we have produced $x_i,y_i\in A_i$ with opposite colours, so $A_i$ is not monochromatic. \qed

\noindent\textbf{5) VERIFICATION}\par
\begin{itemize}
\item Lemma 602.1 does not use the intersection hypotheses; it works for any countable family of infinite sets.
\item Lemma 602.2 isolates the obstruction for uncountable families: at stage $i$ we need at least two previously uncoloured points in $A_i$, which is guaranteed if $A_i$ contributes infinitely many new elements beyond earlier sets.
\item In the original problem, only pairwise intersections are finite. For uncountable $I$, the union $\bigcup_{j<i}(A_i\cap A_j)$ can be infinite, so the ``infinitely many new points'' hypothesis of Lemma 602.2 need not follow.
\end{itemize}

\noindent\textbf{6) FINAL}\par
\textbf{UNRESOLVED}

(i) \emph{Strongest proved partial result.} If the index set $I$ is countable, then such a 2-colouring exists (Lemma 602.1). More generally, if the family can be well-ordered so that each $A_i$ has infinitely many new points outside the union of earlier sets, then a 2-colouring exists (Lemma 602.2).

(ii) \emph{First gap (crisp).} For an uncountable family $(A_i)_{i\in I}$ with $|A_i|=\aleph_0$ and pairwise intersections finite and $\neq 1$, prove or refute the existence of a 2-colouring of $\cup A_i$ with no monochromatic $A_i$.

(iii) \emph{Top 3 next moves.}
\begin{enumerate}
\item Attempt to show that the condition ``$|A_i\cap A_j|\neq 1$'' forces the existence of a well-order of $I$ satisfying the ``infinitely many new points'' hypothesis, or some weaker variant sufficient for a transfinite greedy colouring.
\item Search for a concrete uncountable candidate counterexample where every $A_i$ is ``almost covered'' by intersections with earlier sets in any well-order, so the greedy method cannot pick fresh points.
\item Finite-model sanity: experiment with finite approximations (finite sets with the same intersection pattern) to see which intersection patterns tend to obstruct 2-colourings, and whether excluding intersection size $1$ removes those obstructions.
\end{enumerate}

(iv) \emph{What a minimal counterexample would look like.} A minimal counterexample would be an uncountable index set $I$ and countable sets $(A_i)$ with pairwise intersections finite and never of size $1$, such that for every 2-colouring $c$ of $\cup A_i$ there exists some $i\in I$ with $c$ constant on $A_i$. Structurally, such a family would have to defeat every attempt to assign two opposite-coloured ``fresh'' points per $A_i$.
