
OUTPUT
1) FORMAL RESTATEMENT
Let $Q_n$ be the $n$-dimensional hypercube graph with vertex set $\{0,1\}^n$ and edges between
vertices at Hamming distance $1$. Let
\[f(n)=\max\{|E(G)|: G\subseteq Q_n\text{ and }G\text{ contains no (not-necessarily-induced) }C_4\}.\]
The conjecture is that
\[f(n)\le \left(\tfrac12+o(1)\right)n2^{n-1} \quad (n\to\infty).\]

2) QUICK LITERATURE/CONTEXT CHECK
Browsing is not available. I only record what the problem statement itself claims about known
bounds; these are not re-proved here.

3) ATTACK PLAN
Proof strategies:
1. Characterize $C_4$ in $Q_n$ as a 2-dimensional face, then use counting or averaging to
   bound the number of edges if all faces are missing an edge.
2. Construct explicit $C_4$-free subgraphs with many edges to obtain lower bounds.

Disproof strategies:
1. Build a $C_4$-free subgraph with density strictly above $1/2$ for all large $n$.

Chosen path: provide a clean lower bound construction ($\ge \tfrac12 e(Q_n)$) and the basic
structural lemma about 2-faces.

4) WORK
Lemma 1 (C4 = 2-face in the hypercube).
Every 4-cycle in $Q_n$ lies in a unique 2-dimensional subcube (square face), and every 2-face is a
4-cycle.

Proof.
A length-4 cycle flips exactly two coordinates, each twice, so its vertices are
$x, x\oplus e_i, x\oplus e_j, x\oplus e_i\oplus e_j$ for some distinct $i,j$. These form a 2-face.
Conversely, fixing all coordinates except $i,j$ yields exactly this 4-cycle. ∎

Lemma 2 (Half-layer construction).
Let $H\subseteq Q_n$ be the subgraph containing exactly those edges whose higher-weight endpoint
has odd Hamming weight. Then $H$ is $C_4$-free and
\[|E(H)|=\tfrac12\,|E(Q_n)|=n2^{n-2}.\]

Proof.
Any 2-face has vertices of weights $r,r+1,r+1,r+2$. Its four edges are split between the layers
$L_{r+1}$ and $L_{r+2}$, which have opposite parity. Thus $H$ contains at most two edges from any
face, so no $C_4$ appears (Lemma 1). For the edge count, the number of edges between levels
$k-1$ and $k$ is $k\binom{n}{k}$. Summing over odd $k$ gives
\[\sum_{k\text{ odd}} k\binom{n}{k} = \tfrac12\sum_{k} k\binom{n}{k} = \tfrac12 n2^{n-1}=n2^{n-2},\]
using the standard binomial identity. ∎

Corollary 3 (Elementary lower bound).
For all $n\ge 2$, $f(n)\ge n2^{n-2}=\tfrac12 e(Q_n)$.

Proof.
Immediate from Lemma 2. ∎

5) VERIFICATION
Edge cases:
- $n=1$: $Q_1$ has one edge and no $C_4$.
- $n=2$: $Q_2$ is a 4-cycle; the construction keeps exactly two edges.

Quantifier checks:
Lemma 2 yields a $C_4$-free subgraph for every $n$, so the lower bound holds uniformly.

6) FINAL
**UNRESOLVED**

(i) Strongest fully proved partial result:
$f(n)\ge \tfrac12 e(Q_n)$ via the half-layer construction (Lemma 2).

(ii) First gap:
An upper bound forcing any $C_4$-free subgraph to have at most $(\tfrac12+o(1))e(Q_n)$ edges.

(iii) Top 3 next moves:
1. Count 2-faces and apply extremal inequalities to relate missing edges to face counts.
2. Develop a stability result: if $|E(G)|>(1/2+\delta)e(Q_n)$ then some face must be complete.
3. Explore algebraic constructions to test whether density $>1/2$ is achievable without $C_4$.

(iv) Minimal counterexample structure (if the conjecture is false):
A sequence of $C_4$-free subgraphs with density bounded away from $1/2$ and with a structured
pattern that avoids full 2-faces.
