% Attempt for Erdos Problem #30
% Following PROMPT_STRATEGY.MD
% Tools/Constraints:
% - Web browsing available? NO (restricted)
% - Computation available? YES (local Python), not used
% GPT 5.2

OUTPUT
1) FORMAL RESTATEMENT
A set $A\subseteq\{1,\dots,N\}$ is a \emph{Sidon set} (a $B_2$-set) if whenever
$a,b,c,d\in A$ and $a+b=c+d$, then $\{a,b\}=\{c,d\}$.
Let $h(N)$ be the maximum possible size of a Sidon subset of $\{1,\dots,N\}$.
Question (as stated): is it true that for every $\varepsilon>0$,
\[h(N)=N^{1/2}+O_\varepsilon(N^\varepsilon)?\]

2) QUICK LITERATURE/CONTEXT CHECK
Browsing is not available. I only record what the problem statement itself claims.
I have not verified those results here.

3) ATTACK PLAN
Proof track:
1. Use the distinct-differences characterization to get upper bounds via counting available differences.
2. Sharpen the counting argument using refined structure of sumsets/differences.

Disproof track:
1. Construct Sidon sets via finite geometry (Singer construction) to match the lower bound.
2. Try to force larger fluctuations in $h(N)-\sqrt N$.

Chosen path: prove a clean elementary upper bound $h(N)\le \sqrt{2N}+1$.

4) WORK
Lemma 1 (Sidon $\Rightarrow$ distinct positive differences).
Let $A=\{a_1<\cdots<a_m\}\subseteq\{1,\dots,N\}$ be Sidon. Then the positive differences
\[a_j-a_i\quad (1\le i<j\le m)
\]
are all distinct.

Proof.
Suppose $a_j-a_i=a_{j'}-a_{i'}$ with $i<j$ and $i'<j'$. Then
$a_j+a_{i'}=a_{j'}+a_i$. By the Sidon property, $\{a_j,a_{i'}\}=\{a_{j'},a_i\}$.
Since $a_j>a_i$ and $a_{j'}>a_{i'}$, the only possibility is $a_j=a_{j'}$ and $a_i=a_{i'}$, hence $(i,j)=(i',j')$.
Thus differences are distinct. \qed

Lemma 2 (Difference counting bound).
If $A\subseteq\{1,\dots,N\}$ is Sidon with $|A|=m$, then
\[\binom{m}{2}\le N-1.
\]

Proof.
By Lemma 1 the $\binom{m}{2}$ differences $a_j-a_i$ are distinct positive integers.
All such differences lie in $\{1,2,\dots,N-1\}$, which has size $N-1$. \qed

Corollary 3 (Elementary upper bound).
For all $N\ge 1$,
\[h(N)\le \frac{1+\sqrt{1+8(N-1)}}{2} < \sqrt{2N}+1.
\]

Proof.
Apply Lemma 2 with $m=h(N)$ and solve the quadratic inequality $m(m-1)/2\le N-1$. \qed

5) VERIFICATION
- Lemma 1 uses only the Sidon definition and a case check on set equality.
- Lemma 2 is a direct pigeonhole argument.

6) FINAL
**UNRESOLVED**

(i) Strongest fully proved partial result:
A self-contained upper bound $h(N)<\sqrt{2N}+1$ (Corollary 3).

(ii) First gap:
No method here approaches the conjectured error term $O_\varepsilon(N^\varepsilon)$ or even $O(N^{1/4})$.

(iii) Top 3 next moves:
1. Reproduce the classical $N^{1/2}+O(N^{1/4})$ bound in a self-contained way and track constants.
2. Investigate stability: if $|A|$ is close to $\sqrt N$, what structure must $A$ have?
3. Study Singer-type constructions for sharp lower bounds and compare to the conjectured error term.

(iv) Minimal counterexample structure (if the conjecture were false):
A sequence $N_j$ for which every maximal Sidon set has size $\sqrt{N_j}+\omega(N_j^\varepsilon)$ for some fixed $\varepsilon>0$,
forcing persistent large deviations from $\sqrt N$.
