% Attempt for Erdos Problem #32
% Following PROMPT_STRATEGY.MD
% Tools/Constraints:
% - Web browsing available? NO (restricted)
% - Computation available? YES (local Python), not used
% GPT 5.2

OUTPUT
1) FORMAL RESTATEMENT
A set $A\subseteq\mathbb N$ is an \emph{additive complement to the primes} if every sufficiently large integer $n$
can be written as
\[n=p+a\quad\text{with $p$ prime and $a\in A$.}\]
Questions (as stated):
(a) Can one choose such an $A$ with $|A\cap[1,N]|=o((\log N)^2)$?
(b) Can one achieve $|A\cap[1,N]|=O(\log N)$?
(c) Must any such $A$ satisfy $\liminf_{N\to\infty} |A\cap[1,N]|/\log N >1$?

2) QUICK LITERATURE/CONTEXT CHECK
Browsing is not available. I only record what the problem statement itself claims.
I have not verified those results here.

3) ATTACK PLAN
Proof track:
1. Use probabilistic method: choose $A$ randomly with density about $(\log N)^2/N$ on dyadic intervals and show coverage.
2. Prove lower bounds on $|A\cap[1,N]|$ by a counting argument using $\pi(N)$.

Disproof track:
1. Search for congruence obstructions preventing $O(\log N)$ complements.

Chosen path: prove an unconditional counting lower bound in terms of $\pi(N)$.

4) WORK
Lemma 1 (Counting lower bound via prime counting).
Assume every integer $n$ with $1\le n\le N$ can be written as $n=p+a$ with $p$ prime and $a\in A$.
Let $m(N):=|A\cap[1,N]|$ and let $\pi(N)$ be the number of primes $\le N$. Then
\[m(N)\,\pi(N)\ge N.
\]

Proof.
For each $a\in A\cap[1,N]$, the integers representable as $p+a$ with $p\le N$ form a set of size at most $\pi(N)$.
Thus the total number of integers in $[1,N]$ that can be covered using shifts by elements of $A\cap[1,N]$ is at most
$m(N)\pi(N)$ (union bound). Since by hypothesis all $N$ integers are covered, we get $m(N)\pi(N)\ge N$. \qed

Corollary 2 (Conditional logarithmic lower bound).
Assuming the prime number theorem estimate $\pi(N)\sim N/\log N$, Lemma 1 implies
\[m(N)\ge (1+o(1))\log N.
\]

Proof.
From Lemma 1, $m(N)\ge N/\pi(N)$; substitute $\pi(N)\sim N/\log N$. \qed

5) VERIFICATION
- Lemma 1 is a coarse union bound and ignores overlaps; it is still valid.
- Corollary 2 depends on the prime number theorem, which is not proved here.

6) FINAL
**UNRESOLVED**

(i) Strongest fully proved partial result:
A necessary condition for covering $[1,N]$ is $|A\cap[1,N]|\ge N/\pi(N)$ (Lemma 1).

(ii) First gap:
No construction here achieves $|A\cap[1,N]|=o((\log N)^2)$, and no unconditional lower bound of order $\log N$ is proved
without using deep estimates for $\pi(N)$.

(iii) Top 3 next moves:
1. Implement Erd\H{o}s's probabilistic construction (as described in the problem statement) in a self-contained way.
2. Strengthen Lemma 1 by incorporating overlaps (e.g. via inclusion--exclusion or sieve) to approach the constant $e^\gamma$ bound.
3. Investigate whether $O(\log N)$ is possible by building explicit near-optimal complements.

(iv) Minimal counterexample structure (if $O(\log N)$ were impossible):
An obstruction showing that for infinitely many $N$, the shifts $\mathcal P + a$ (for $a\in A\cap[1,N]$) cannot cover
$[1,N]$ unless $|A\cap[1,N]|$ exceeds $C\log N$ for some $C>0$.
