% Attempt for Erdos Problem #38
% Following PROMPT_STRATEGY.MD
% Tools/Constraints:
% - Web browsing available? NO (restricted)
% - Computation available? YES (local Python), not used
% GPT 5.2

OUTPUT
1) FORMAL RESTATEMENT
Definitions.
- The Schnirelmann density of $A\subseteq\mathbb N$ is
\[d_s(A):=\inf_{N\ge 1} \frac{|A\cap\{1,\dots,N\}|}{N}.\]
So $d_s(A)=\alpha$ means $|A\cap[1,N]|\ge \alpha N$ for every $N$ and the inequality is essentially sharp for some $N$.
- For $b\in\mathbb N$, write $A+b:=\{a+b:a\in A\}$.
- "$B$ is an additive basis" (as used in additive number theory) typically means: there exists $k\ge 1$ such that every
sufficiently large integer is a sum of $k$ (not necessarily distinct) elements of $B$.

Question.
Does there exist a set $B\subset\mathbb N$ that is not an additive basis, and a function
$f:(0,1)\to (0,\infty)$, such that for every $\alpha\in(0,1)$, every $A\subseteq\mathbb N$ with $d_s(A)=\alpha$, and every
$N\ge 1$, there exists $b\in B$ with
\[ |(A\cup(A+b))\cap[1,N]| \ge (\alpha+f(\alpha))N? \]

2) QUICK LITERATURE/CONTEXT CHECK
Web browsing is not available in this session. I only record what is stated in the problem text:
- Erd\H{o}s proved the inequality with gain $\alpha(1-\alpha)/(2k)$ when $B$ is an additive basis of order $k$.
- There are connections to "essential components" and to work of Linnik.
I have not verified any of these external claims.

3) ATTACK PLAN
Proof track:
1. Derive necessary conditions on $B$ and on $f(\alpha)$ by choosing special sets $A$ with controlled translation behavior.
2. Try to leverage known structure theorems for Schnirelmann density and sumsets to build (or rule out) such a $B$.

Disproof track:
1. Find an obstruction: a family of sets $A$ (with fixed Schnirelmann density) such that for all $b$ in a candidate $B$,
   the union $A\cup(A+b)$ cannot gain a positive proportion.
2. In particular, test periodic sets $A$ and use gcd/parity obstructions for $B$.

Chosen path: obtain elementary necessary conditions on $f$ and $B$ (fully proved).

4) WORK
Lemma 1 (Trivial upper bound on $f$).
For any $\alpha\in(0,1)$, any $A\subseteq\mathbb N$, any $b\in\mathbb N$, and any $N\ge 1$, one has
\[|(A\cup(A+b))\cap[1,N]|\le N.\]
Consequently, any function $f$ appearing in the problem must satisfy
\[f(\alpha)\le 1-\alpha\qquad(0<\alpha<1).\]

Proof.
The set $(A\cup(A+b))\cap[1,N]$ is a subset of $[1,N]$, so its cardinality is at most $N$.
If the problem's inequality holds, then for each $\alpha$ there exist choices with
$(\alpha+f(\alpha))N\le N$, i.e. $f(\alpha)\le 1-\alpha$. \qed

Lemma 2 (If the property holds, then $\gcd(B)=1$).
Assume $B\subset\mathbb N$ satisfies the stated property for some function $f$ with $f(\alpha)>0$ for all $\alpha\in(0,1)$.
Then $\gcd(B)=1$.

Proof.
Suppose for contradiction that $g:=\gcd(B)\ge 2$. Then every $b\in B$ is a multiple of $g$.
Define
\[A:=\{n\in\mathbb N: n\not\equiv 0\pmod g\}.\]
Claim: $d_s(A)=1-\frac{1}{g}$. Indeed, for any $N\ge 1$ we have
\[|A\cap[1,N]| = N-\left\lfloor\frac{N}{g}\right\rfloor \ge N-\frac{N}{g} = \left(1-\frac{1}{g}\right)N,\]
so $d_s(A)\ge 1-1/g$. For $N=g$ we have $|A\cap[1,g]|=g-1$, hence
\[\frac{|A\cap[1,g]|}{g}=1-\frac{1}{g},\]
so $d_s(A)\le 1-1/g$. This proves the claim.

Now fix any $b\in B$. Since $b\equiv 0\pmod g$, translation by $b$ preserves the residue class $0\pmod g$.
Therefore $n\equiv 0\pmod g$ iff $n-b\equiv 0\pmod g$, which implies $A+b=A$.
Hence for every $N$,
\[|(A\cup(A+b))\cap[1,N]| = |A\cap[1,N]|.\]
In particular, for $N$ a multiple of $g$ we have $|A\cap[1,N]|=(1-1/g)N$.
Let $\alpha:=d_s(A)=1-1/g\in(0,1)$. The stated property would require that for this $A$ and this $N$ there exists
$b\in B$ such that
\[|(A\cup(A+b))\cap[1,N]|\ge (\alpha+f(\alpha))N.\]
But the left-hand side equals $\alpha N$, so this inequality forces $f(\alpha)\le 0$, contradicting $f(\alpha)>0$.
Thus $g=1$.
\qed

5) VERIFICATION
- Lemma 1 is a pure set-theoretic bound and is exact.
- Lemma 2 uses an explicit periodic test set $A$ and computes its Schnirelmann density exactly.
  The contradiction uses $N$ a multiple of $g$ so that $|A\cap[1,N]|=(1-1/g)N$.

6) FINAL
**UNRESOLVED**

(i) Strongest fully proved partial result:
Any admissible gain function must satisfy $f(\alpha)\le 1-\alpha$, and any $B$ with the stated property must have $\gcd(B)=1$
(Lemmas 1--2).

(ii) First gap:
Either (a) construct a specific $B$ with $\gcd(B)=1$ that is not an additive basis but still forces a uniform Schnirelmann-density gain,
or (b) prove that no such $B$ can exist even when $\gcd(B)=1$.

(iii) Top 3 next moves (concrete):
1. Try to strengthen Lemma 2: for any fixed finite set of residues $R\subset\mathbb Z/m\mathbb Z$, analyze sets $A$ that are unions
   of residue classes and quantify the best possible gain under translation by $b\in B$.
2. Search for further necessary conditions on $B$ by choosing $A$ as complements of sparse periodic sets (to test many $\alpha$).
3. Attempt constructive candidates for $B$ based on known "essential components" that are not additive bases, and check whether they
   can enforce a uniform gain $f(\alpha)>0$.

(iv) What a minimal counterexample would likely look like (if existence fails):
A set $B$ with $\gcd(B)=1$ but so arithmetically constrained that one can build, for some $\alpha\in(0,1)$, a periodic (or almost periodic)
set $A$ of Schnirelmann density $\alpha$ whose translates by all $b\in B$ overlap $A$ too heavily to yield any positive proportional gain.
