\section{Problem 106971: Homotopy type of $\mathrm{TOP}(4)/\mathrm{PL}(4)$}

\subsection*{1) FORMAL RESTATEMENT}
\paragraph{Literal statement.}
The attachment asks whether $\mathrm{TOP}(4)/\mathrm{PL}(4)$ is homotopy equivalent to $K(\Z/2,3)$, given that the natural map
\[
\mathrm{TOP}(4)/\mathrm{PL}(4)\longrightarrow \mathrm{TOP}/\mathrm{PL}\simeq K(\Z/2,3)
\]
is $5$-connected.

\paragraph{Corrected proposition.}
\[
\mathrm{TOP}(4)/\mathrm{PL}(4)\simeq K(\Z/2,3)\quad\text{(homotopy equivalence?)}.
\]

\subsection*{2) QUICK LITERATURE/CONTEXT CHECK (web available)}
I did not locate, during the allotted browsing, a definitive published resolution of whether the above $5$-connected map is a homotopy equivalence.

\subsection*{3) ATTACK PLAN}
\begin{itemize}
\item \textbf{Proof track:} Show higher $\pi_k$ vanish for $k\ge 6$ (e.g.\ via smoothing/stabilization).
\item \textbf{Disproof track:} Exhibit nontrivial higher $\pi_k$ for some $k\ge 6$.
\end{itemize}

\subsection*{4) WORK (proved consequence of ``5-connected'')}
Assume as given (from the attachment) that the map
\[
f:\mathrm{TOP}(4)/\mathrm{PL}(4)\to K(\Z/2,3)
\]
is $5$-connected.

\begin{lemma}
Under this hypothesis,
\[
\pi_1=\pi_2=0,\quad \pi_3\cong \Z/2,\quad \pi_4=\pi_5=0
\]
for $\mathrm{TOP}(4)/\mathrm{PL}(4)$.
\end{lemma}

\begin{proof}
A $5$-connected map induces isomorphisms on $\pi_i$ for $i\le 4$ and a surjection on $\pi_5$.
Since $K(\Z/2,3)$ has $\pi_3=\Z/2$ and $\pi_i=0$ for $i\neq 3$, the conclusion follows.
\end{proof}

\subsection*{5) VERIFICATION}
The lemma is correct assuming $5$-connectedness. The open part is whether $\pi_k=0$ for $k\ge 6$.

\subsection*{6) FINAL}
\textbf{UNRESOLVED.}

\medskip
\noindent\textbf{Fail-safe details.}
\begin{enumerate}[label=(\roman*)]
\item \textbf{Strongest fully proved partial result:} The low-degree homotopy computation conditional on $5$-connectedness.
\item \textbf{First gap:} Determine $\pi_k$ for $k\ge 6$ (or show $f$ is a weak equivalence).
\item \textbf{Top 3 next moves:} Postnikov analysis; family smoothing obstructions; stabilization comparison with $\mathrm{TOP}(n)/\mathrm{PL}(n)$ for $n\ge 5$.
\item \textbf{Minimal counterexample:} A nontrivial element in some $\pi_k$ for $k\ge 6$ or a nontrivial higher Postnikov invariant.
\end{enumerate}
