\section{Problem 2: $f\circ f=g$ revisited (linear subspaces of compositional squares)}

\subsection*{1) FORMAL RESTATEMENT}
Let $C(\R)$ be the real vector space of all continuous functions $\R\to\R$ with pointwise operations.
Define
\[
\mathcal S := \{g\in C(\R):\ \exists f\in C(\R)\text{ with } f\circ f = g\}.
\]
Question: does there exist a linear subspace $V\subset C(\R)$ with $\dim V\ge 2$ (or $\dim V=\infty$) such that $V\subset \mathcal S$?

\medskip
\noindent\textbf{Stress points.}
\begin{itemize}[leftmargin=2em]
\item The property ``$g$ has a continuous iterative square root'' is highly nonlinear and not stable under addition/scalar multiplication.
\item If $V$ is a linear subspace of $C(\R)$ contained in $\mathcal S$, then $g\in V\Rightarrow -g\in V\subset\mathcal S$. Thus every $g\in V$ must satisfy both $g\in\mathcal S$ and $-g\in\mathcal S$; this is a strong symmetry constraint.
\end{itemize}

\subsection*{2) QUICK LITERATURE/CONTEXT CHECK (web)}
The MO page contains comments noting (i) $\mathcal S$ contains no \emph{decreasing} functions (cf.\ a related MO question on $f(f(x))=g(x)$), and (ii) any linear subspace $V\subset\mathcal S$ must contain no injective functions (since injective continuous functions on $\R$ are monotone). As of \date{}, the MO page shows no posted answer.

\subsection*{3) ATTACK PLAN}
\textbf{Proof-track candidates (nonexistence).}
\begin{enumerate}[label=(P\arabic*),leftmargin=2.5em]
\item Show that if $g\in\mathcal S$ and $g$ is ``sufficiently nondegenerate'' (e.g.\ injective, or locally monotone on an unbounded set), then $-g\notin\mathcal S$. Conclude any $V\subset\mathcal S$ is extremely constrained.
\item Try to prove: if $g\in\mathcal S\cap(-\mathcal S)$ then $g$ must be constant, which would imply $\dim V\le 1$.
\end{enumerate}
\textbf{Construction-track candidates (existence).}
\begin{enumerate}[label=(D\arabic*),leftmargin=2.5em]
\item Find an explicit infinite-dimensional linear subspace $V$ of ``highly non-injective'' functions and build continuous square roots for all $g\in V$ by a uniform coding/decoding construction.
\item Find a linear operator $L:C(\R)\to C(\R)$ whose image lies in $\mathcal S$ (then $V=\mathrm{im}(L)$ is a linear subspace of $\mathcal S$).
\end{enumerate}
Best available progress here is (P1): a complete obstruction for injective functions.

\subsection*{4) WORK}

\paragraph{PHASE 1: tiny cases by hand.}
\begin{itemize}[leftmargin=2em]
\item Constant functions: for $g(x)\equiv c$, take $f(x)\equiv c$, so $g\in\mathcal S$.
Thus the 1-dimensional subspace $\{x\mapsto c:\ c\in\R\}\subset \mathcal S$.
\item Translations: $g(x)=x+t$ has square root $f(x)=x+t/2$, so $g\in\mathcal S$.
But $-g(x)=-x-t$ is strictly decreasing and (as shown below) cannot be in $\mathcal S$.
So translations cannot lie in any linear subspace $V\subset\mathcal S$ (except $\{0\}$).
\end{itemize}

\paragraph{Lemma 1 (injective functions cannot lie in a linear subspace of $\mathcal S$).}
Let $g\in C(\R)$ be injective and suppose $g\in\mathcal S$.
Then $-g\notin\mathcal S$.
Consequently, no linear subspace $V\subset C(\R)$ with $V\subset\mathcal S$ can contain a nonzero injective function.

\begin{proof}
Assume $g\in\mathcal S$, so $\exists f\in C(\R)$ with $f\circ f=g$.
\emph{Step 1: $f$ is injective.} If $f(x)=f(y)$, then applying $f$ gives $g(x)=f(f(x))=f(f(y))=g(y)$.
Since $g$ is injective, $x=y$. Hence $f$ is injective.

\emph{Step 2: $f$ is strictly monotone.} A continuous injective map $f:\R\to\R$ is strictly monotone (standard real analysis: otherwise it would take the same value twice by the intermediate value theorem).

\emph{Step 3: $g$ is strictly increasing.} If $f$ is strictly increasing, then $f\circ f$ is strictly increasing.
If $f$ is strictly decreasing, then $f\circ f$ is strictly increasing as a composition of two decreasing maps.
Thus in either case $g=f\circ f$ is strictly increasing.

Now consider $-g$. Since $g$ is strictly increasing and injective, $-g$ is strictly decreasing and injective.

Suppose for contradiction that $-g\in\mathcal S$. Then $\exists h\in C(\R)$ with $h\circ h=-g$.
As above, since $-g$ is injective, $h$ is injective, hence strictly monotone.
But then $h\circ h$ is strictly increasing (composition of two monotone maps of the same monotonicity).
This contradicts the fact that $-g$ is strictly decreasing.
Therefore $-g\notin\mathcal S$.
The final claim follows because a linear subspace $V$ satisfies $g\in V\Rightarrow -g\in V$.
\end{proof}

\paragraph{Strongest fully proved partial result obtained here.}
Lemma~1 is a complete obstruction: any linear subspace $V\subset\mathcal S$ contains no nonzero injective functions (in particular, no nonzero monotone functions).

\paragraph{First hard gap.}
I do not know whether there exists any nonconstant $g\in\mathcal S$ with $-g\in\mathcal S$.
Proving that $\mathcal S\cap(-\mathcal S)$ consists only of constants would resolve the MO question in the negative.
Constructing an explicit nonconstant element of $\mathcal S\cap(-\mathcal S)$ (or better, an infinite-dimensional subspace) would resolve it in the positive.

\subsection*{5) VERIFICATION}
\begin{itemize}[leftmargin=2em]
\item Lemma~1 hinges on two facts: (i) if $g$ is injective and $g=f\circ f$, then $f$ is injective; (ii) continuous injective $\R\to\R$ implies strict monotonicity. Both are standard and were justified.
\item The step ``$h\circ h$ is increasing'' is checked in both cases: if $h$ increasing, composition increases; if $h$ decreasing, composition increases.
\end{itemize}

\subsection*{6) FINAL}
\textbf{UNRESOLVED.}

\noindent (i) Strongest partial: Lemma~1 (no injective elements in any $V\subset\mathcal S$).

\noindent (ii) First gap: determine whether $\mathcal S\cap(-\mathcal S)$ contains a nonconstant function.

\noindent (iii) Top 3 next moves:
\begin{enumerate}[leftmargin=2.5em]
\item Attempt to prove $\mathcal S\cap(-\mathcal S)$ consists only of constants by extracting a dynamical obstruction (e.g.\ monotonicity on invariant sets, existence of monotone branches, periodic orbit structure) that cannot be symmetric under $g\mapsto -g$ unless $g$ is constant.
\item Try to construct a nonconstant $g$ with $g,-g\in\mathcal S$ by a piecewise-defined non-injective $f$ whose second iterate is sign-flexible.
\item Search for a linear operator $L$ with $\mathrm{im}(L)\subset\mathcal S$ using coding on disjoint intervals and ensuring the induced condition is linear.
\end{enumerate}

\noindent (iv) Likely minimal counterexample structure (if $V$ exists): a space of functions all constant on large sets (hence highly non-injective), avoiding creation of strictly monotone functions under linear combinations.
