\section{Problem 3: Minimal volume of certain 4-manifolds and smooth structures}

\subsection*{1) FORMAL RESTATEMENT}
For $k\in\N$, define the closed oriented topological 4-manifold
\[
M_k := \#^{\,k}(S^2\times S^2)\ \#\ \#^{\,1+k}(S^1\times S^3),
\]
where $\#$ denotes connected sum.

For a smooth closed $n$-manifold $X$, define its \emph{minimal volume}
\[
\mathrm{MinVol}(X) := \inf\big\{\mathrm{Vol}(X,g):\ g\ \text{Riemannian metric on }X,\ \abs{K_g}\le 1\ \text{everywhere}\big\},
\]
where $K_g$ denotes sectional curvature.

The (yes/no) question is:
\begin{quote}
For $k$ sufficiently large, suppose $X$ is a smooth manifold homeomorphic to $M_k$ with $\mathrm{MinVol}(X)=0$. Must $X$ be diffeomorphic to $M_k$ with its standard smooth structure?
\end{quote}
Equivalently: do all exotic smooth structures on the topological manifold $M_k$ have $\mathrm{MinVol}>0$?

\medskip
\noindent\textbf{Stress points.}
\begin{itemize}[leftmargin=2em]
\item $\mathrm{MinVol}(X)=0$ means there exist metrics with uniformly bounded sectional curvature and volumes $\to 0$ (no diameter control is assumed).
\item The Cheeger--Gromov theory relates $\mathrm{MinVol}=0$ to existence of (polarized) $F$-structures, but the converse direction requires care.
\item The question is sensitive to smooth structure: Kotschick produced homeomorphic but nondiffeomorphic 4-manifolds with different minimal volume behavior.
\end{itemize}

\subsection*{2) QUICK LITERATURE/CONTEXT CHECK (web)}
The MO post attributes to Kotschick the existence, for large $k$, of distinct smooth structures on $M_k$ such that the standard one has $\mathrm{MinVol}=0$ (via a fixed-point-free circle action) while another has $\mathrm{MinVol}>0$. The question asks whether $\mathrm{MinVol}=0$ forces the smooth structure to be standard. As of \date{}, the MO page shows no posted answer.

\subsection*{3) ATTACK PLAN}
\textbf{Proof-track candidates (show $\mathrm{MinVol}=0\Rightarrow$ standard).}
\begin{enumerate}[label=(P\arabic*),leftmargin=2.5em]
\item Use collapse theory: try to show $\mathrm{MinVol}=0$ implies existence of an $F$-structure of positive rank, then classify $F$-structures (or circle actions) on manifolds in this homeomorphism type and compare with known classification of circle actions (Fintushel) and Seiberg--Witten constraints (Baldridge).
\item Extract smooth-structure-sensitive invariants forced to vanish under bounded-curvature collapse (e.g.\ Seiberg--Witten invariants, Yamabe invariant, entropy); then check whether exotic structures on $M_k$ violate these vanishing constraints.
\end{enumerate}
\textbf{Disproof-track candidates (find exotic $X$ with $\mathrm{MinVol}(X)=0$).}
\begin{enumerate}[label=(D\arabic*),leftmargin=2.5em]
\item Construct an exotic smooth structure on $M_k$ that nonetheless admits an $F$-structure (or fixed-point-free $S^1$-action), yielding $\mathrm{MinVol}=0$.
\item Use known constructions of exotic 4-manifolds via surgeries that preserve collapse structures.
\end{enumerate}

\subsection*{4) WORK}

\paragraph{PHASE 0/1: basic invariants and sanity checks.}
Compute Betti numbers for $M_k$:
\[
b_1(M_k) = 1+k,\qquad b_2(M_k)=2k,\qquad b_+(M_k)=k,\qquad \chi(M_k)= 2k.
\]
(Here each $S^2\times S^2$ contributes $b_2=2$ and $b_+=1$, and each $S^1\times S^3$ contributes $b_1=1$; connected sums add Betti numbers in these degrees.)

In particular $b_+(M_k)=k$, so for $k\ge 2$ we are in the $b_+>1$ regime where Seiberg--Witten invariants are defined without chamber ambiguities.

\paragraph{PHASE 1: what is easy to prove from the definition.}
The standard smooth manifold $M_k$ has $\mathrm{MinVol}(M_k)=0$ because it admits a fixed-point-free $S^1$-action (as stated in the attachment), hence a polarized $F$-structure, which implies bounded-curvature collapse by the Cheeger--Gromov construction.
(Here I treat the Cheeger--Gromov implication ``polarized $F$-structure $\Rightarrow \mathrm{MinVol}=0$'' as a known theorem; reproducing its full proof would be far beyond the scope of this writeup.)

\paragraph{Strongest fully proved partial result obtained here.}
A purely topological consequence: $\mathrm{MinVol}(M_k)=0$ for the standard smooth structure for all $k$ (given the existence of a fixed-point-free circle action/F-structure).

\paragraph{First hard gap.}
The key missing step for a \emph{proof} is a converse: from $\mathrm{MinVol}(X)=0$ infer a sufficiently rigid geometric/topological structure (like an $F$-structure or an actual smooth circle action) that is known to constrain smooth structures on $M_k$.
I do not know how to make that inference in general, and I do not know a construction of an exotic smooth $X\cong_{\mathrm{homeo}} M_k$ with $\mathrm{MinVol}(X)=0$.

\subsection*{5) VERIFICATION}
\begin{itemize}[leftmargin=2em]
\item The Betti number computation is standard and can be checked directly by Mayer--Vietoris for connected sums.
\item The implication ``fixed-point-free $S^1$-action $\Rightarrow \mathrm{MinVol}=0$'' is not reproved here; it is a known result via $F$-structures and collapse theory.
\end{itemize}

\subsection*{6) FINAL}
\textbf{UNRESOLVED.}

\noindent (i) Strongest partial: identification of the basic invariants and the (known) implication that the standard structure has $\mathrm{MinVol}=0$.

\noindent (ii) First gap: no established route from $\mathrm{MinVol}=0$ to rigidity of smooth structure in this homeomorphism type.

\noindent (iii) Top 3 next moves:
\begin{enumerate}[leftmargin=2.5em]
\item Check whether existing ``collapse with bounded curvature $\Rightarrow F$-structure'' theorems apply without diameter bounds in this dimension, possibly after rescaling/normalization.
\item Investigate whether $\mathrm{MinVol}=0$ forces vanishing of Seiberg--Witten invariants (or related smooth invariants) for $b_+>1$ in dimension 4; compare with Kotschick's exotic examples.
\item Attempt to build an exotic smooth structure on $M_k$ that still supports an $F$-structure by adapting known surgery constructions that preserve torus actions or $F$-structures.
\end{enumerate}

\noindent (iv) Likely minimal counterexample structure (if the answer is ``no''): an exotic smooth manifold homeomorphic to $M_k$ admitting an $F$-structure (or at least a sequence of bounded-curvature collapsing metrics) yet not diffeomorphic to the standard connected sum.
