\section{Greatly expanded new edition of a Bourbaki chapter on algebra? (MO 228330)}

\subsection*{1) FORMAL RESTATEMENT}
\paragraph{Ambiguity.} ``Written recently'' and ``revival'' are vague. The post asks a historical/factual question rather than a mathematical proposition.

\paragraph{Minimal corrected statement (falsifiable).}
We interpret the question as the conjunction of two precise claims:
\begin{enumerate}[leftmargin=2em]
\item[(C1)] The 2012 Springer edition ``Alg\`ebre: Chapitre 8'' is a substantially revised (``entirely rewritten'') edition of the 1958 text, not a simple reprint.
\item[(C2)] Bourbaki's \emph{\`El\'ements de math\'ematique} project continued to produce new/revised volumes after 2012.
\end{enumerate}

\subsection*{2) QUICK LITERATURE/CONTEXT CHECK}
Springer lists ``Alg\`ebre, Chapitre 8'' with copyright 2012 and ``Topologie alg\'ebrique, Chapitres 1 \`a 4'' with copyright 2016 in the \emph{\`El\'ements de math\'ematique} series listing. \href{https://www.springer.com/series/7436/books}{(Springer series page)}.

A bibliographic listing describes the 2012 French edition of Chapter 8 as a ``deuxi\`eme \`edition enti\`erement refondue de l'\'edition de 1958'' with publication date 2012-01-14 and 489 pages. \href{https://www.barnesandnoble.com/w/algebre-n-bourbaki/1149063688}{(Barnes \& Noble listing)}.

\subsection*{3) ATTACK PLAN}
\begin{itemize}[leftmargin=2em]
\item Prove (C1) by citing a bibliographic source that explicitly states the 2012 edition is an entirely rewritten second edition of the 1958 edition.
\item Prove (C2) by exhibiting at least one post-2012 volume in the Bourbaki \emph{\`El\'ements} series with a publication/copyright date after 2012.
\end{itemize}

\subsection*{4) WORK}
\begin{theorem}[Evidence for (C1) and (C2)]
Claims (C1) and (C2) hold.
\end{theorem}

\begin{proof}
\textbf{(C1)} The Barnes \& Noble bibliographic entry for ``Alg\`ebre: Chapitre 8'' (Springer) states explicitly: ``Ce volume est une deuxi\`eme \`edition enti\`erement refondue de l'\'edition de 1958'' and gives publication date 01/14/2012 and 489 pages. This directly supports that the 2012 edition is an ``entirely reworked'' second edition of the 1958 edition.

\textbf{(C2)} Springer maintains a page listing book titles in the \emph{\`El\'ements de math\'ematique} series. On that list, ``Topologie alg\'ebrique, Chapitres 1 \`a 4'' is listed with copyright 2016 (and available renditions), while ``Alg\`ebre, Chapitre 8'' is listed with copyright 2012. The presence of a 2016 volume in the same series establishes that the project produced volumes after 2012.
\end{proof}

\subsection*{5) VERIFICATION}
\paragraph{Adversarial checks.}
The argument does not claim anything about \emph{who} wrote the revision or the internal Bourbaki process. It only establishes the two factual claims (C1)--(C2) using bibliographic/series evidence.

\subsection*{6) FINAL}
\textbf{PROOF.}
