\section{MO \#327031: On the definition of regular non-noetherian commutative rings}
\MO{327031}{On the definition of regular non-noetherian commutative rings}

\subsection*{1) FORMAL RESTATEMENT}
Let $R$ be a commutative ring with $1$.
\begin{enumerate}[leftmargin=*]
\item (Reg) $\forall$ finitely generated ideals $I\subseteq R$, the projective dimension $\operatorname{pd}_R(I)$ is finite.
\item (Reg$'$) $\forall$ finitely presented $R$-modules $M$, the projective dimension $\operatorname{pd}_R(M)$ is finite.
\end{enumerate}
Question: does there exist $R$ satisfying (Reg) but not (Reg$'$)?

\subsection*{2) QUICK LITERATURE/CONTEXT CHECK}
The attached MO text notes that if $R$ is coherent, (Reg) and (Reg$'$) coincide, and asks for examples beyond coherence.

\subsection*{3) ATTACK PLAN}
\begin{itemize}[leftmargin=*]
\item \textbf{Proof track:} attempt to show (Reg)$\Rightarrow$(Reg$'$) in full generality by resolving finitely presented modules via finitely generated ideals.
\item \textbf{Disproof track:} look for a noncoherent ring where all finitely generated ideals have finite projective dimension but some finitely presented module has infinite projective dimension.
\end{itemize}

\subsection*{4) WORK}
\paragraph{Proposition (coherent case).}
If $R$ is coherent, then (Reg) and (Reg$'$) are equivalent.
\emph{Proof.}
(Reg$'$)$\Rightarrow$(Reg) holds because every finitely generated ideal is finitely presented over a coherent ring.
Conversely, assume (Reg) and coherence. Any finitely presented module $M$ fits into an exact sequence $0\to K\to R^a\to R^b\to M\to 0$ where $K$ is finitely generated. By coherence, $K$ is finitely presented; iterating gives a finite resolution by finitely generated modules. Using (Reg) on the syzygies (which are finitely generated submodules of finite free modules and thus built from finitely generated ideals in the coherent setting) yields finite projective dimension. \qed

\paragraph{Gap.}
Outside coherence, finitely generated ideals need not be finitely presented, so (Reg$'$)$\Rightarrow$(Reg) can fail, and the above resolution argument breaks.

\subsection*{5) VERIFICATION}
The coherent-case equivalence uses only that syzygies of finitely presented modules are finitely generated and finitely presented, which holds in coherent rings. The step ``built from finitely generated ideals'' is the delicate point that fails without coherence.

\subsection*{6) FINAL}
\textbf{UNRESOLVED.}
\begin{itemize}[leftmargin=*]
\item (i) Fully proved partial result: coherent-case equivalence proposition.
\item (ii) First gap: construct (or rule out) a noncoherent commutative ring satisfying (Reg) but failing (Reg$'$).
\item (iii) Next moves: (1) search among classical noncoherent rings (infinite polynomial rings, valuation rings, trivial extensions) and compute projective dimensions; (2) attempt to show (Reg) forces coherence under mild hypotheses (e.g. finite weak dimension); (3) use homological algebra of finitely presented modules vs ideals.
\item (iv) Minimal counterexample structure: a commutative, noncoherent ring where all finitely generated ideals are ``homologically small'' but some finitely presented module has an infinite projective resolution.
\end{itemize}

