\section{MO 363120: ``Players alternate moving a $\{\swarrow,\uparrow,\rightarrow\}$ piece on a chessboard''}
\label{sec:mo363120}
\noindent\textbf{MathOverflow link:} \url{https://mathoverflow.net/questions/363120/players-alternate-moving-a-swarrow-uparrow-rightarrow-piece-on-a-chessb} (accessed 2026-01-16).

\subsection*{1) FORMAL RESTATEMENT}
Fix $k\in\NN$. Consider the board
\[
B_k := \{0,1,\dots,2k\}\times\{0,1,\dots,2k\} \subset \ZZ^2.
\]
A single token starts at a specified square $(x_0,y_0)\in B_k$.
Two players alternate moves; a move from $(x,y)$ may go to any of:
\[
(x-1,y-1),\quad (x,y+1),\quad (x+1,y),
\]
provided the destination lies in $B_k$ and has not been visited previously by the token.
The first player who cannot move loses (normal play).

\medskip
\noindent\textbf{Question.}
Describe the set of losing starting squares (for the first player), i.e.
\[
L_k := \{(x,y)\in B_k: \text{the first player loses with optimal play starting at }(x,y)\}.
\]
The MO post reports experimental evidence that all even--even squares (both coordinates even) are losing, with a small number of additional exceptions.

\subsection*{2) QUICK LITERATURE/CONTEXT CHECK}
The MathOverflow page has no posted answers as of 2026-01-16.
The post mentions an equivalent formulation in terms of a divisor game on divisors of $10^{2k}$ (reachable positions correspond to divisors and moves correspond to multiplying/dividing by $2,5,10$), but no general solution is provided.

\subsection*{3) ATTACK PLAN}
\textbf{Proof-track ideas.}
\begin{enumerate}[leftmargin=2em]
\item Try to find a pairing/mirroring strategy proving that a large class of starts are losing (e.g. all even--even).
\item Seek an invariant or monotone potential to show that certain starts are winning by forcing a move to a losing class.
\end{enumerate}

\textbf{Disproof-track ideas.}
\begin{enumerate}[leftmargin=2em]
\item Compute $L_k$ for small $k$ by exact minimax recursion (state includes visited set). Compare with the conjectured description.
\item Attempt to produce a $k$ where some even--even start is actually winning.
\end{enumerate}

We carried out the exact minimax computation for $k=1,2,3$.

\subsection*{4) WORK}
\subsubsection*{4.1 Exact computation for small boards}
We solved the game by recursion on the state $(\text{current square},\ \text{visited set})$ with memoization.
For each start square, visited set initially contains only the start.
A state is winning iff there exists a legal move to a losing state.

\medskip
\noindent\textbf{Results.}
Write coordinates as $(x,y)$ with $0\le x,y\le 2k$.
\begin{itemize}[leftmargin=2em]
\item For $k=1$ (board $3\times3$):
\[
L_1 = \{(0,0),(0,2),(2,0),(2,2)\}.
\]
These are exactly the even--even squares.
\item For $k=2$ (board $5\times5$): $L_2$ consists of the $9$ even--even squares $(x,y)$ with $x,y\in\{0,2,4\}$.
\item For $k=3$ (board $7\times7$): $L_3$ contains all $16$ even--even squares $(x,y)$ with $x,y\in\{0,2,4,6\}$, and additionally
\[
(3,4),(4,5),(5,0),(5,6).
\]
\end{itemize}
Thus the heuristic ``even--even squares are losing'' is consistent with these computations, but it is not a complete description once $k\ge 3$.

\subsubsection*{4.2 Partial structural observations}
The move set contains directed $3$-cycles
\[
(x,y)\to(x+1,y)\to(x+1,y+1)\to(x,y),
\]
so the underlying directed graph is not acyclic; the no-revisit constraint is essential.
The game is a special instance of \emph{vertex geography} on a directed graph, a class known to be computationally hard in general.

\subsection*{5) VERIFICATION}
\begin{itemize}[leftmargin=2em]
\item Computation was exact for $k\le 3$ (state spaces: $9,25,49$ vertices; still manageable under memoization).
\item No general theorem was proved; no claim is made beyond the computed values.
\item The additional losing squares for $k=3$ show that ``even--even iff losing'' is false in general.
\end{itemize}

\subsection*{6) FINAL}
\begin{center}
\textbf{UNRESOLVED}
\end{center}


