\section{MO 377545: ``Which finite abelian groups aren't homotopy groups of spheres?''}
\label{sec:mo377545}
\noindent\textbf{MathOverflow link:} \url{https://mathoverflow.net/questions/377545/which-finite-abelian-groups-arent-homotopy-groups-of-spheres} (accessed 2026-01-16).

\subsection*{1) FORMAL RESTATEMENT}
\textbf{Literal issue.} The MO post asks an open-ended classification question.
We extract a concrete yes/no sub-question explicitly asked in the post.

\medskip
\noindent\textbf{Concrete subproblem.}
Does there exist a pair $(n,k)$ with $n\ge 2$, $k>n$ such that
\[
\pi_k(S^n) \cong \ZZ/5\ZZ\ ?
\]
More broadly, characterize the finite abelian groups that occur as $\pi_k(S^n)$.

\subsection*{2) QUICK LITERATURE/CONTEXT CHECK}
The MathOverflow page has no posted answers as of 2026-01-16.
From standard background:
\begin{itemize}[leftmargin=2em]
\item For $k>n\ge 2$, $\pi_k(S^n)$ is finite (Serre).
\item $5$-torsion does occur in low-dimensional homotopy groups, e.g. tables indicate groups like $\pi_{10}(S^3)\cong \ZZ/15\ZZ$ and $\pi_{10}(S^2)\cong \ZZ/15\ZZ$ (hence containing elements of order $5$), but this does not answer whether a group is \emph{exactly} $\ZZ/5\ZZ$.
\end{itemize}

\subsection*{3) ATTACK PLAN}
\textbf{Proof-track ideas.}
\begin{enumerate}[leftmargin=2em]
\item Search known tables for an explicit instance with $\pi_k(S^n)\cong\ZZ/5$.
\item Use structural results (EHP sequence, unstable Adams spectral sequence) to prove existence/nonexistence of isolated $5$-torsion.
\end{enumerate}

\textbf{Disproof-track ideas.}
\begin{enumerate}[leftmargin=2em]
\item Prove that whenever $5$-torsion appears in $\pi_k(S^n)$, additional torsion (e.g. $3$-torsion or $2$-torsion) necessarily accompanies it, ruling out $\ZZ/5$.
\end{enumerate}

We did not obtain either an explicit $(n,k)$ with $\ZZ/5$ or a theorem excluding it.

\subsection*{4) WORK}
\subsubsection*{4.1 Small-case checks from published tables}
Published tables (e.g. Wikipedia compilations of low-dimensional homotopy groups) show $5$-torsion in groups such as $\pi_{10}(S^3)$ and $\pi_{10}(S^2)$, but those groups are $\ZZ/15$ rather than $\ZZ/5$.
No example of a homotopy group equal to $\ZZ/5$ was located in these low-dimensional tables in this writeup.

\subsubsection*{4.2 A fully proved partial fact: finiteness for $k>n\ge 2$}
\textbf{Proposition.} If $n\ge 2$ and $k>n$, then $\pi_k(S^n)$ is finite.

\textbf{Proof sketch.}
This is a classical theorem of Serre: using the Serre spectral sequence for the path-loop fibration and properties of homology with finite coefficients, one proves that the homotopy groups of simply-connected spheres are finite in the stable range $k>n$.
(We do not reproduce Serre's full proof here.) \qed

\subsection*{5) VERIFICATION}
\begin{itemize}[leftmargin=2em]
\item The proposition is standard but invoked as a named theorem; details are not reproduced.
\item No definitive progress was made on the existence of $\ZZ/5$ as an entire homotopy group.
\end{itemize}

\subsection*{6) FINAL}
\begin{center}
\textbf{UNRESOLVED}
\end{center}


