\section{MO 467258 --- A conjecture about inclusion--exclusion}
\label{sec:mo467258}
\noindent\textbf{MathOverflow link:} \href{https://mathoverflow.net/questions/467258}{https://mathoverflow.net/questions/467258}.

\subsection*{1) FORMAL RESTATEMENT}
Fix $n\in\mathbb{N}$. Let $S_1,\dots,S_n$ be subsets of a universe $U$.
Assume they are pairwise incomparable: for $i\ne j$, neither $S_i\subseteq S_j$ nor $S_j\subseteq S_i$.

For each nonempty $T\subseteq\{1,\dots,n\}$ write
\[
S_T := \bigcap_{i\in T} S_i.
\]
In the inclusion--exclusion expansion of $\bigcup_i S_i$, when different $T$ yield equal intersections $S_T$, their coefficients may cancel.
Call an intersection set $X$ \emph{non-cancelling} if the total coefficient of $X$ after merging equal terms is nonzero.

\medskip
\textbf{Conjecture (as in the MO post).}
The union $\bigcup_i S_i$ can be constructed from the collection of non-cancelling intersections using only:
\begin{itemize}
  \item disjoint union, and
  \item relative complement $A\setminus B$ when $B\subseteq A$.
\end{itemize}

\subsection*{2) QUICK LITERATURE/CONTEXT CHECK}
The MO post references a short note by the same author: arXiv:2401.16210, which proves special cases and provides further discussion.
As of writing, the MO question appears unanswered.

\subsection*{3) ATTACK PLAN}
\textbf{Proof strategies.}
\begin{itemize}
  \item Model intersections as elements of a finite poset (by inclusion) and attempt to express the union via a disjoint ``atom'' decomposition using only intersections with nonzero M\"obius coefficients.
  \item Try induction on the size of the family or on the number of distinct intersections.
\end{itemize}
\textbf{Disproof strategies.}
\begin{itemize}
  \item Search for smallest $n$ with a complicated intersection-poset where a non-cancelling intersection is ``missing'' an atom needed to build the union.
  \item Build examples where many distinct $S_T$ coincide and coefficients cancel massively.
\end{itemize}

\subsection*{4) WORK}
\textbf{PHASE 1 (tiny cases).}
\begin{itemize}
  \item $n=1$ trivial.
  \item $n=2$ has no cancellation under the incomparability hypothesis; the union is built from $S_1,S_2,S_{12}$ by $(S_1\setminus S_{12})\sqcup (S_2\setminus S_{12})\sqcup S_{12}$.
\end{itemize}

\medskip
\textbf{What is currently missing.}
For $n\ge 3$, equality among distinct intersections $S_T$ can occur even with incomparability of singletons, and the conjecture demands that \emph{only} those intersections with nonzero net coefficient be used as building blocks.
A complete proof or counterexample was not obtained here.

\subsection*{5) VERIFICATION}
The $n=2$ decomposition is checked and uses only allowed operations.
No overclaims are made for $n\ge 3$.

\subsection*{6) FINAL}
\textbf{UNRESOLVED}.
\begin{itemize}
  \item (i) Partial result: full verification for $n=2$ under incomparability.
  \item (ii) First gap: handle cancellation patterns for $n\ge 3$.
  \item (iii) Next moves: (1) formalize the ``intersection poset'' and M\"obius inversion viewpoint; (2) attempt explicit constructive algorithm using nonzero M\"obius terms; (3) computer search over finite universes for smallest counterexample.
  \item (iv) A plausible minimal counterexample would have $n=3$ or $4$ with several distinct $T$ giving the same $S_T$, cancelling a mid-level intersection that seems needed to carve out Venn atoms.
\end{itemize}

% -----------------------------------------------------------------------------
