\section{MO 478499 --- Is this generalized character always a character?}
\label{sec:mo478499}
\noindent\textbf{MathOverflow link:} \href{https://mathoverflow.net/questions/478499}{https://mathoverflow.net/questions/478499}.

\subsection*{1) FORMAL RESTATEMENT}
Let $G$ be a finite group and $p$ a prime.
Define a class function $\Psi\colon G\to\mathbb{Z}$ by:
\begin{itemize}
  \item If $p\mid |g|$ then $\Psi(g)=0$.
  \item If $g$ is a $p'$-element (i.e. $p\nmid |g|$), then
  \[
    \Psi(g) := |\{x\in C_G(g): x \text{ has order a power of }p\}|.
  \]
\end{itemize}
It is known (and assumed) that $\Psi$ is a generalized character.
\textbf{Question.} Must $\Psi$ be a genuine character (a nonnegative integer combination of irreducible characters)?

\subsection*{2) QUICK LITERATURE/CONTEXT CHECK}
The MO post itself notes:
\begin{itemize}
  \item in groups of Lie type in characteristic $p$, $\Psi$ agrees with the square of the Steinberg character;
  \item sufficient conditions involving normal $p$-complements in centralizers;
  \item an update (Nov 2025) pointing to a paper by G.~Robinson discussing this in a stronger form.
\end{itemize}

\subsection*{3) ATTACK PLAN}
\textbf{Proof track.}
\begin{itemize}
  \item Show $\langle \Psi,\chi\rangle\ge 0$ for every irreducible $\chi$.
  \item Use structural reduction: normal $p$-complements, block theory, or groups of Lie type.
\end{itemize}
\textbf{Disproof track.}
\begin{itemize}
  \item Search for smallest group $G$ and prime $p$ where some inner product $\langle\Psi,\chi\rangle$ is negative.
\end{itemize}

\subsection*{4) WORK}
\textbf{PHASE 1 (tiny sanity).}
\begin{itemize}
  \item If $p\nmid |G|$, then every element is $p'$ and the only $p$-power element is $1$, so $\Psi\equiv 1$ is the trivial character.
  \item If $G$ is a $p$-group, then $\Psi(1)=|G|$ and $\Psi(g)=0$ for $g\ne 1$, so $\Psi$ is the regular character.
\end{itemize}

\medskip
\textbf{What remains open here.}
The general question ``always a character?'' was not proved or disproved in this writeup.

\subsection*{5) VERIFICATION}
The two extreme cases above are correct and give genuine characters.

\subsection*{6) FINAL}
\textbf{UNRESOLVED}.
\begin{itemize}
  \item (i) Partial results: verification in the cases $p\nmid |G|$ and $G$ a $p$-group.
  \item (ii) First gap: prove nonnegativity of all character inner products in general (or exhibit a group/prime with a negative one).
  \item (iii) Next moves: (1) implement a computational search over small groups using character tables; (2) use Robinson's criteria (centralizer normal $p$-complements); (3) test families like alternating groups and small groups of Lie type outside defining characteristic.
  \item (iv) A minimal counterexample (if it exists) would likely be a small non-solvable group with mixed $p$-local structure where $\Psi$ has oscillating values on $p'$-classes.
\end{itemize}

% -----------------------------------------------------------------------------

% -----------------------------------------------------------------------------
