\section{MO 68094: Ricci-flat metric on $n$-sphere}

\subsection*{1) FORMAL RESTATEMENT}
\paragraph{Literal statement.}
For an integer $n>4$, does there exist a smooth Riemannian metric $g$ on the smooth manifold $S^n$ such that $\mathrm{Ric}(g)=0$ everywhere?

\paragraph{Conventions.}
$S^n$ denotes the smooth $n$-sphere. ``Ricci-flat'' means the Ricci curvature tensor vanishes identically.

\subsection*{2) QUICK LITERATURE/CONTEXT CHECK}
The MathOverflow thread contains no answer and includes a comment indicating the problem is open for $n>4$ (and in fact already for $n=4$), along with a comment by R.\ Bryant noting that any metric on $S^n$ has holonomy $SO(n)$.\footnote{MO page \url{https://mathoverflow.net/questions/68094/ricci-flat-metric-on-n-sphere}.}
The same page links to another MO question about existence of compact simply connected Ricci-flat manifolds with generic holonomy.

\subsection*{3) ATTACK PLAN}
\paragraph{Proof-track.}
Try to construct a Ricci-flat metric on $S^n$ via known Ricci-flat construction methods (Calabi--Yau, $G_2$, $\mathrm{Spin}(7)$, gluing). Immediate obstruction: these methods produce reduced holonomy, which is incompatible with $S^n$'s cohomology.

\paragraph{Disproof-track.}
Try to find a topological obstruction to Ricci-flatness on spheres (Bochner technique, index theory, special holonomy). Works in low dimensions ($n=2,3$), but for $n\ge 4$ no known general obstruction is standard.

\paragraph{Best path chosen.}
Prove rigorously the nonexistence for $n=2$ and $n=3$ (sanity checks), and isolate the open case $n>4$.

\subsection*{4) WORK}
\subsubsection*{Phase 1: tiny cases}
\paragraph{$n=2$.}
On a surface, $\mathrm{Ric}=Kg$ where $K$ is Gauss curvature. Ricci-flat means $K\equiv 0$. By Gauss--Bonnet,
\[
\int_{S^2} K\,dA = 2\pi\chi(S^2)=4\pi\neq 0,
\]
contradiction. Hence no Ricci-flat metric on $S^2$.

\paragraph{$n=3$.}
In dimension $3$, the Riemann curvature tensor is determined by the Ricci tensor via:
\[
R_{ijkl}= g_{ik}R_{jl}+g_{jl}R_{ik}-g_{il}R_{jk}-g_{jk}R_{il}
-\frac{R}{2}(g_{ik}g_{jl}-g_{il}g_{jk}),
\]
where $R_{jl}$ is Ricci and $R$ is scalar curvature. If $\mathrm{Ric}=0$, then $R=0$ and the right-hand side is $0$, so the full curvature tensor vanishes: the metric is flat. A closed flat $3$-manifold has infinite fundamental group (its universal cover is $\mathbb{R}^3$), but $\pi_1(S^3)=0$, contradiction. Hence no Ricci-flat metric on $S^3$.

\subsubsection*{Phase 2: $n>4$ landscape}
If $S^n$ carried a Ricci-flat metric and were simply connected, Berger's holonomy theorem implies the holonomy is one of $SO(n)$ or special groups ($SU(m)$, $Sp(k)$, $G_2$, $\mathrm{Spin}(7)$). Special holonomy would force existence of a nonzero parallel differential form of degree $k$ with $0<k<n$; such a form is harmonic and hence defines a nontrivial cohomology class. But $H^k(S^n)=0$ for $0<k<n$, ruling out those special holonomies. Thus any hypothetical Ricci-flat metric on $S^n$ must have full holonomy $SO(n)$, consistent with Bryant's comment.

No general existence or obstruction result for such a metric on $S^n$ (for $n\ge 4$) is supplied in the available sources.

\subsection*{5) VERIFICATION}
\begin{itemize}[leftmargin=2em]
\item The $n=2$ obstruction uses only Gauss--Bonnet (fully standard).
\item The $n=3$ argument uses the explicit curvature decomposition in dimension $3$ plus the classification of closed flat manifolds (Bieberbach): a flat closed manifold has $\mathbb{R}^n$ universal cover and hence infinite $\pi_1$.
\item For $n>4$, we verified that special-holonomy Ricci-flat is impossible on $S^n$ due to cohomology, but this does not exclude full holonomy $SO(n)$.
\end{itemize}

\subsection*{6) FINAL}
\begin{center}
\textbf{UNRESOLVED}
\end{center}

\paragraph{(i) Strongest proved partial result here.}
No Ricci-flat metrics exist on $S^2$ or $S^3$. Any Ricci-flat metric on $S^n$ for $n\ge 4$ would necessarily have full holonomy $SO(n)$ (no special holonomy).

\paragraph{(ii) First gap.}
Existence or obstruction for Ricci-flat metrics with full holonomy on $S^n$ for $n>4$.

\paragraph{(iii) Top 3 next moves.}
(1) Investigate gluing constructions that can yield Ricci-flat metrics with generic holonomy on simply connected manifolds; (2) attempt to derive a topological obstruction specific to spheres (e.g.\ via index-theoretic invariants, stable homotopy, or moduli of Einstein metrics); (3) analyze whether any Ricci-flat metric on a homology sphere must be flat (likely false in high dimension, but worth targeting).

\paragraph{(iv) Minimal counterexample.}
An explicit Ricci-flat metric on some $S^n$ ($n>4$), necessarily with holonomy $SO(n)$.
