\section{MO 69819: Three-dimensional hyperbolic volume of a four-manifold}

\subsection*{1) FORMAL RESTATEMENT}
\paragraph{Ambiguity/misstatement.}
In the quoted definition, the link $L$ lies in $\#_g(S^2\times S^1)$ but the complement is written $(S^2\times S^1)\setminus L$; the minimal correction is to replace it by $(\#_g(S^2\times S^1))\setminus L$.

\paragraph{Corrected definition.}
Let $X$ be a smooth closed orientable $4$-manifold and fix a handle decomposition. Let $H_{0,1}$ denote the union of the $0$- and $1$-handles; then $\partial H_{0,1}\cong \#_g(S^2\times S^1)$. The attaching circles of the $2$-handles form a framed link $L\subset \#_g(S^2\times S^1)$.

Define the \emph{3D volume} of the handle decomposition to be the simplicial volume (Gromov norm)
\[
\mathrm{Vol}_3(\text{decomp}) := \left\| \#_g(S^2\times S^1)\setminus L \right\|
\]
of the link complement (with torus boundary). By Geometrization, this equals (up to the universal constant $v_3$) the sum of volumes of the hyperbolic pieces in the JSJ decomposition.

Define the \emph{3D volume} of $X$ to be
\[
\mathrm{Vol}_3(X):=\inf_{\text{handle decompositions of }X}\ \mathrm{Vol}_3(\text{decomp}).
\]
The MO post asserts the infimum is a minimum because the set of hyperbolic volumes is well-ordered (J\o rgensen--Thurston).

\paragraph{Questions.}
Classify $X$ with $\mathrm{Vol}_3(X)=0$; decide whether $\mathrm{Vol}_3(X)>0$ can occur.

\subsection*{2) QUICK LITERATURE/CONTEXT CHECK}
The MO page has no posted answers; it includes a comment suggesting a relation to F-structures and vanishing simplicial/minimal volume, and a reference to Gromov--Guth for related ideas.\footnote{MO page \url{https://mathoverflow.net/questions/69819/what-is-the-three-dimensional-hyperbolic-volume-of-a-four-manifold}.}
The well-ordering of volumes of hyperbolic 3-manifolds is a classical theorem of J\o rgensen--Thurston.\footnote{See e.g.\ expository references such as the StackExchange discussion with citations: \url{https://mathoverflow.net/questions/220889/is-there-an-infinite-strictly-decreasing-sequence-of-volumes-of-hyperbolic-3-manifolds}, and standard 3-manifold texts.}
(Here we only use the qualitative fact ``well-ordered'' to justify minimum vs infimum, not in computations.)

\subsection*{3) ATTACK PLAN}
\paragraph{Landscape / tools.}
\begin{itemize}[leftmargin=2em]
\item Kirby calculus / handle slides: changes link $L$ without changing $X$.
\item JSJ decomposition and simplicial volume: $\|M\|=0$ iff hyperbolic pieces absent.
\item Graph manifolds: precisely the prime 3-manifolds all of whose JSJ pieces are Seifert fibered.
\item F-structures / collapse theory (Cheeger--Gromov): often imply simplicial volume and minimal volume vanish.
\item Try explicit families: doubles of 2-handlebodies, simply connected manifolds, aspherical 4-manifolds.
\end{itemize}

\paragraph{Proof-track.}
Prove large classes have $\mathrm{Vol}_3(X)=0$ by exhibiting handle decompositions with graph-manifold link complements.

\paragraph{Disproof-track.}
Find a 4-manifold $X$ such that every handle decomposition necessarily has a link complement with a hyperbolic JSJ piece (positive simplicial volume). This would give $\mathrm{Vol}_3(X)>0$.

\paragraph{Best path chosen.}
Prove rigorously the easy examples of $\mathrm{Vol}_3(X)=0$ mentioned in the MO post; identify the first gap for positivity.

\subsection*{4) WORK}
\subsubsection*{Phase 1: sanity checks / examples with volume $0$}
\paragraph{Example 1: $S^4$.}
Use the trivial handle decomposition with one $0$-handle and one $4$-handle. Then $g=0$ and $L=\varnothing\subset S^3$. The complement is $S^3$, which has simplicial volume $0$. Hence $\mathrm{Vol}_3(S^4)=0$.

\paragraph{Example 2: $\mathbb{C}P^2$.}
A standard handle decomposition of $\mathbb{C}P^2$ has a $0$-handle, a single $2$-handle attached along an unknot in $S^3$ with framing $+1$, and a $4$-handle. Here $g=0$ and the attaching link complement is $S^3\setminus$ unknot, which is a solid torus, hence a Seifert fibered (graph) manifold and has simplicial volume $0$. Therefore $\mathrm{Vol}_3(\mathbb{C}P^2)=0$.

\paragraph{Example 3: doubles of 2-handlebodies as in the MO post.}
Let $W$ be a compact 4-manifold built from a 0-handle and 2-handles attached along a framed link $L\subset \#_g(S^2\times S^1)$, and assume $\#_g(S^2\times S^1)\setminus L$ is a graph manifold. Then the handle decomposition of $W$ has 3D volume $0$. Doubling $W$ along $\partial W$ yields a closed 4-manifold $X=\mathrm{double}(W)$; the same 0-1-2 handle description exhibits $X$ with 3D volume $0$. Hence such doubles satisfy $\mathrm{Vol}_3(X)=0$.

\subsubsection*{Phase 2: what is not proved}
The difficult question is whether there exists any $X$ with $\mathrm{Vol}_3(X)>0$.
A comment on the MO page suggests: if $\mathrm{Vol}_3(X)=0$, then perhaps $X$ admits an F-structure and hence has vanishing simplicial volume (and likely minimal volume $0$), but this is only stated as a suspicion, not a proof.\footnote{See Ian Agol's comment on the MO page \url{https://mathoverflow.net/questions/69819/what-is-the-three-dimensional-hyperbolic-volume-of-a-four-manifold}.}

\subsection*{5) VERIFICATION}
\begin{itemize}[leftmargin=2em]
\item The examples use explicit handle decompositions where the attaching link complement is a graph manifold. Since graph manifolds have simplicial volume $0$, these decompositions have 3D volume $0$ by definition.
\item None of the arguments addresses whether \emph{every} handle decomposition of a given $X$ must have positive volume; this is the core difficulty for positivity.
\end{itemize}

\subsection*{6) FINAL}
\begin{center}
\textbf{UNRESOLVED}
\end{center}

\paragraph{(i) Strongest fully proved partial result here.}
Many standard 4-manifolds (e.g.\ $S^4$, $\mathbb{C}P^2$, and doubles of suitable 2-handlebodies) have $\mathrm{Vol}_3(X)=0$ because they admit handle decompositions with graph-manifold attaching link complements.

\paragraph{(ii) Exact first gap.}
Produce a single explicit closed 4-manifold $X$ and prove that \emph{every} handle decomposition has $\#_g(S^2\times S^1)\setminus L$ with positive simplicial volume (i.e.\ contains a hyperbolic JSJ piece), thereby forcing $\mathrm{Vol}_3(X)>0$.

\paragraph{(iii) Top 3 next moves.}
(1) Relate $\mathrm{Vol}_3(X)$ to known 4D invariants (simplicial volume of $X$, minimal volume, presence of F-structures); (2) examine aspherical 4-manifolds with large simplicial volume and attempt to obstruct graph-manifold level sets; (3) study the effect of Kirby moves on simplicial volume of link complements to see whether hyperbolic volume can always be eliminated.

\paragraph{(iv) Minimal counterexample.}
A 4-manifold with a proven obstruction to having graph-manifold link complements in \emph{all} handle decompositions (e.g.\ arising from fundamental group or simplicial volume constraints).
