\section{MO 73428: When is a compact topological $4$-manifold a CW complex?}

\subsection*{1) FORMAL RESTATEMENT}

\paragraph{Conventions.}
A \emph{topological $4$-manifold} means a Hausdorff, second-countable topological space $M$
such that every point has a neighborhood homeomorphic to $\mathbb{R}^4$ (or to $\mathbb{R}^4_+$ in the presence of boundary).
For a \emph{compact} manifold, second-countability is automatic.
A \emph{CW complex} means a Hausdorff space obtained by attaching open cells $e^n_\alpha\cong\mathbb{R}^n$
in the usual inductive way; a \emph{finite} CW complex has finitely many cells. Any compact CW complex is necessarily finite.

\paragraph{Ambiguity / misstatement.}
The prompt is a \emph{question} (“When is ...?”), not a proposition with truth value. So there is no literal statement to prove/disprove.

\paragraph{Minimal corrected statement.}
The mathematically standard sharpening compatible with the question is:

\medskip
\noindent\textbf{Corrected Problem.}
\emph{Characterize (or determine whether there exist) compact topological $4$-manifolds $M$ that are \textbf{not} homeomorphic to a finite CW complex.}
Equivalently: \emph{Is every compact topological $4$-manifold homeomorphic to a finite CW complex?}
\medskip

\paragraph{Stress points / edge cases.}
\begin{itemize}[leftmargin=2em]
\item Boundary vs.\ closed: $M$ may or may not have boundary.
\item “CW complex” as a \emph{homeomorphism type} is much stronger than “has the homotopy type of a CW complex”.
\item Compactness forces finiteness for CW complexes, so the question is really about \emph{finite} CW structures.
\item Known nontriangulable compact $4$-manifolds (e.g.\ Freedman’s $E_8$ manifold) are natural test cases.
\end{itemize}

\subsection*{2) QUICK LITERATURE/CONTEXT CHECK (web available)}

A modern survey notes the following (paraphrased precisely):
\begin{itemize}[leftmargin=2em]
\item For every compact topological $n$-manifold with $n\ge 5$, there is a finite CW complex $X$ such that $M$ is homeomorphic to the \emph{mapping cylinder} of a map $\partial M\to X$ (and in particular for closed $n\ge 5$, $M$ is homeomorphic to a finite CW complex).
\item The analogous question in dimension $4$ (even for closed manifolds) is explicitly listed as open.
\end{itemize}
(See Theorem 4.2 and Question 4.4 in the cited survey.)

A separate MathOverflow thread states that closed topological manifolds of dimension $\ge 6$ are homeomorphic to CW complexes and discusses the status in low dimensions.

\subsection*{3) ATTACK PLAN}

\paragraph{Proof-track strategies.}
\begin{enumerate}[leftmargin=2em]
\item \textbf{Resolution/cell-like approximation approach.}
Show any compact topological $4$-manifold is the image of a cell-like map from a finite polyhedron, then attempt to upgrade to a homeomorphism by shrinking arguments.
\item \textbf{Controlled topology approach.}
Use controlled handle decompositions / end-theorems to build a CW filtration directly in dimension $4$.
\item \textbf{ANR recognition plus extra structure.}
Manifolds are ANRs; try to show compact $4$-manifolds satisfy additional local tameness implying CW structure.
\end{enumerate}

\paragraph{Disproof-track strategies.}
\begin{enumerate}[leftmargin=2em]
\item \textbf{Look for an invariant obstructing “being a finite CW complex” as a homeomorphism type.}
For example, properties of local contractibility plus combinatorial tameness conditions.
\item \textbf{Try known “bad” 4-manifolds:} Freedman $E_8$, non-smoothable or non-triangulable examples, and try to show they cannot admit any CW structure.
\item \textbf{Non-resolvable ANR-manifold phenomena.}
Construct a compact $4$-manifold as a cell-like image of a polyhedron in a way that provably cannot be a CW complex.
\end{enumerate}

\paragraph{Best path chosen.}
Given the current state of the literature (the problem is listed as open), a full proof/disproof is currently out of reach here.
I will therefore:
\begin{itemize}[leftmargin=2em]
\item Prove the strongest unconditional partial results that can be made gap-free,
\item Identify the exact first gap that blocks a full solution,
\item Describe concrete next moves and what a minimal counterexample would look like.
\end{itemize}

\subsection*{4) WORK (proved partial results)}

\begin{theorem}[Finite CW \emph{homotopy type} for compact manifolds]\label{thm:homotopy-type-finite-CW}
Let $M$ be a compact topological manifold (any dimension). Then $M$ has the homotopy type of a finite CW complex.
\end{theorem}

\begin{proof}
A compact topological manifold is metrizable, finite-dimensional, and locally contractible. In particular it is an ANR (absolute neighborhood retract) in the category of metrizable spaces.
A theorem of West states that every compact ANR has the homotopy type of a finite CW complex.
Applying this theorem to $M$ yields the claim.
\end{proof}

\begin{remark}
Theorem \ref{thm:homotopy-type-finite-CW} addresses the weaker question “does $M$ have the homotopy type of a finite CW complex?” but does \emph{not} imply that $M$ is \emph{homeomorphic} to a CW complex.
\end{remark}

\begin{theorem}[Smoothable implies CW]\label{thm:smoothable-implies-CW}
Let $M$ be a compact smooth $4$-manifold (with or without boundary). Then $M$ is homeomorphic to a finite CW complex.
\end{theorem}

\begin{proof}
Choose a Morse function on $M$ with finitely many critical points.
Morse theory yields a handle decomposition of $M$ with finitely many handles.
Each $k$-handle contributes a $k$-cell in a CW decomposition; attaching maps are the handle attaching maps.
Thus $M$ is homeomorphic to a finite CW complex.
\end{proof}

\begin{remark}
For compact topological $4$-manifolds, Kirby proved: “$M$ admits a (topological) handlebody structure $\iff$ $M$ is smoothable.”
Hence, \emph{within dimension $4$}, “smoothable” is a standard sufficient condition for “CW”.
The open problem is whether there are additional compact (non-smoothable) topological $4$-manifolds that nevertheless admit CW structures, or whether every compact topological $4$-manifold does.
\end{remark}

\subsubsection*{Computational / tiny-case checks}
\begin{itemize}[leftmargin=2em]
\item Standard compact smooth $4$-manifolds ($S^4$, $\mathbb{C}P^2$, $S^2\times S^2$, K3, \dots) satisfy Theorem \ref{thm:smoothable-implies-CW}.
\item The canonical “bad” test case is Freedman’s $E_8$ topological manifold (closed, nontriangulable, non-smoothable). Whether it admits a CW structure is part of the open landscape.
\end{itemize}

\subsection*{5) VERIFICATION (adversarial checks)}

\begin{itemize}[leftmargin=2em]
\item Compactness $\Rightarrow$ any CW structure would be finite. This is a strong global requirement, but does not contradict known invariants of topological manifolds.
\item The ANR $\Rightarrow$ finite CW \emph{homotopy type} theorem is widely used and compatible with all manifold examples; it does not settle homeomorphism.
\item The smoothable $\Rightarrow$ CW argument is standard and robust (Morse/handles).
\item The genuine gap: upgrading “mapping cylinder models / controlled decompositions / ANR structure” to an \emph{actual CW homeomorphism type} in dimension $4$.
\end{itemize}

\subsection*{6) FINAL}

\textbf{UNRESOLVED.}

\paragraph{(i) Strongest proved partial result.}
Every compact topological $4$-manifold has the homotopy type of a finite CW complex (Theorem \ref{thm:homotopy-type-finite-CW}). If it is smoothable, it is homeomorphic to a finite CW complex (Theorem \ref{thm:smoothable-implies-CW}).

\paragraph{(ii) Exact first gap.}
Given a compact topological $4$-manifold $M$, prove or disprove:
\[
\exists\ \text{finite CW complex }X\ \text{and homeomorphism }M\cong X.
\]
Known ANR/homotopy-type results do not upgrade to a homeomorphism classification in dimension $4$.

\paragraph{(iii) Top 3 next moves.}
\begin{enumerate}[leftmargin=2em]
\item Try to construct a CW filtration using controlled topology/handle trading in dimension $4$ (where classical handlebody theory fails without smoothability).
\item Study whether known nontriangulable closed $4$-manifolds admit \emph{cell-like} decompositions that could yield CW structures.
\item Seek an obstruction: identify a topological invariant satisfied by all finite CW complexes but potentially violated by some compact topological $4$-manifolds.
\end{enumerate}

\paragraph{(iv) What a minimal counterexample would likely look like.}
A closed, compact, non-smoothable, non-triangulable topological $4$-manifold with the mildest possible local behavior (still a manifold!) but for which every attempted finite cell decomposition fails—most plausibly something built using exotic $4$-dimensional phenomena (Casson handles, Kirby--Siebenmann class, etc.), with no known obstruction presently.
